### 2024-11-02 final-project.Rmd
data cleaning and documentation

---
title: "final-project"
author: "Murphy John"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(stringr)
library(psych)
```

```{r}
# load data
seer1 <- read.csv("data/caselist_2024.11.csv", header = TRUE)
```

```{r}
# column rename
seer_name <- seer1 %>%
  dplyr::rename(
  age = Age.recode.with..1.year.olds,
  sex = Sex,
  year_diagnosis = Year.of.diagnosis,
  race = Race.ethnicity,
  marriage_status = Marital.status.at.diagnosis,
  tnm2015 = Derived.AJCC.T..7th.ed..2010.2015.,
  tnm2017 = Derived.SEER.Combined.T..2016.2017.,
  tnm2018 = Derived.EOD.2018.T..2018..,
  subtype = Breast.Subtype..2010..,
  grade2017 = Grade.Recode..thru.2017.,
  grade2018 = Derived.Summary.Grade.2018..2018..,
  n2015 = Derived.AJCC.N..7th.ed..2010.2015.,
  n2017 = Derived.SEER.Combined.N..2016.2017.,
  n2018 = Derived.EOD.2018.N..2018..,
  m2015 = Derived.AJCC.M..7th.ed..2010.2015.,
  m2017 = Derived.SEER.Combined.M..2016.2017.,
  m2018 = Derived.EOD.2018.M..2018..,
  radiation = Radiation.recode,
  chemotherapy = Chemotherapy.recode..yes..no.unk.,
  surgery = RX.Summ..Surg.Prim.Site..1998..,
  survival_months = Survival.months,
  survival_status = Vital.status.recode..study.cutoff.used.,
  survival_cause_cancer = SEER.cause.specific.death.classification,
  survival_cause_other = SEER.other.cause.of.death.classification,
  idc_type = Histologic.Type.ICD.O.3
)

# filter
seer_filter <- seer_name %>%
  filter(
    age != "00 years",
    age != "15-19 years",
    sex == "Female",
    survival_months >= 1,
    survival_cause_cancer != "Dead (missing/unknown COD)",
    survival_cause_other != "Dead (attributable to causes other than this cancer dx)",
    grade2017 != "B-cell; pre-B; B-precursor",
    grade2017 != "T-cell"
  )

# clean up the variables
seer_clean <- seer_filter %>%
  mutate(
    # assign race as White, Black, or Other
    race = case_when(
      race == "White" ~ "White",
      race == "Black" ~ "Black",
      TRUE ~ "Other"
    ),
    # assign marriage_status as Yes or No
    marriage_status = case_when(
      marriage_status == "Married (including common law)" ~ "Yes",
      TRUE ~ "No"
    ),

    # assign blank tnm to NA
    tnm2015 = case_when(
      tnm2015 == "Blank(s)" ~ NA_character_,
      TRUE ~ tnm2015
    ),
    tnm2017 = case_when(
      tnm2017 == "Blank(s)" ~ NA_character_,
      TRUE ~ tnm2017
    ),
    tnm2018 = case_when(
      tnm2018 == "Blank(s)" ~ NA_character_,
      TRUE ~ tnm2018
    ),

    # merge tnm to ensure cross-validation
    tnm = coalesce(tnm2015, tnm2017, tnm2018),

    # assign breast cancer subtypes to their more commonly used names
    subtype = case_when(
      subtype == "HR-/HER2-" ~ "Triple Negative",
      subtype == "HR-/HER2+" ~ "HER2 Positive",
      subtype == "HR+/HER2-" ~ "Luminal A",
      subtype == "HR+/HER2+" ~ "Luminal B",
      TRUE ~ NA_character_
    ),
    # assign common grades between 2017 and 2018
    # "1" is well-differentiated, "2" is moderately-differentiated
    # "3" is poorly- differentiated, "4" is undifferentiated
    # see documentation for additional details on assignments
    grade2017 = case_when(
      grade2017 == "Well differentiated; Grade I" ~ "1",
      grade2017 == "Moderately differentiated; Grade II" ~ "2",
      grade2017 == "Poorly differentiated; Grade III" ~ "3",
      grade2017 == "Undifferentiated; anaplastic; Grade IV" ~ "4",
      grade2017 == "H" ~ "3",
      grade2017 == "Unknown" ~ NA_character_,
      TRUE ~ grade2017
    ),
    grade2018 = case_when(
      grade2018 == "A" ~ "1",
      grade2018 == "B" ~ "2",
      grade2018 == "C" ~ "3",
      grade2018 == "D" ~ "4",
      grade2018 == "H" ~ "3",
      grade2018 == "9" ~ NA_character_,
      grade2018 == "Blank(s)" ~ NA_character_,
      TRUE ~ grade2018
    ),
    # then merge the 2017 and 2018 grade columns
    grade = coalesce(grade2017, grade2018),

    # assign N stage type for easier analysis
    n2015 = case_when(
      str_detect(n2015, "N0") ~ "N0",
      str_detect(n2015, "N1|N1N0S") ~ "N1",
      str_detect(n2015, "N2|N2N0S") ~ "N2",
      str_detect(n2015, "N3|N3N0S") ~ "N3",
      str_detect(n2015, "NX") ~ "NX",
      TRUE ~ NA_character_
    ),
    n2017 = case_when(
      str_detect(n2017, "c0|p0") ~ "N0",
      str_detect(n2017, "c1|p1") ~ "N1",
      str_detect(n2017, "c2|p2") ~ "N2",
      str_detect(n2017, "c3|p3") ~ "N3",
      str_detect(n2017, "cX") ~ "NX",
      TRUE ~ NA_character_
    ),
    n2018 = case_when(
      str_detect(n2018, "N0") ~ "N0",
      str_detect(n2018, "N1|N1N0S") ~ "N1",
      str_detect(n2018, "N2|N2N0S") ~ "N2",
      str_detect(n2018, "N3|N3N0S") ~ "N3",
      str_detect(n2018, "NX") ~ "NX",
      TRUE ~ NA_character_
    ),
    # then merge the 2015, 2017, and 2018 N stage columns
    n_stage = coalesce(n2015, n2017, n2018),

     # assign M stage type for easier analysis
    m2015 = case_when(
      str_detect(m2015, "M0") ~ "M0",
      str_detect(m2015, "M1|M1M0S") ~ "M1",
      TRUE ~ NA_character_
    ),
    m2017 = case_when(
      str_detect(m2017, "c0") ~ "M0",
      str_detect(m2017, "c1|p1") ~ "M1",
      TRUE ~ NA_character_
    ),
    m2018 = case_when(
      str_detect(m2018, "M0") ~ "M0",
      str_detect(m2018, "M1|M1M0S") ~ "M1",
      TRUE ~ NA_character_
    ),

    # then merge the 2015, 2017, and 2018 M stage columns
    m_stage = coalesce(m2015, m2017, m2018),

    # simplify the radiation variable
    radiation = case_when(
      str_detect(radiation, "None|Refused|unknown") ~ "No/Unknown",
      TRUE ~ "Yes"
    ),
    surgery = case_when(
      surgery == "0" ~ "No/Unknown",
      surgery == "99" ~ "No/Unknown",
      TRUE ~ "Yes"
    )
    )

# further filter and select only variables used in analysis
seer <- seer_clean %>%
  filter(
    !is.na(subtype)
  ) %>%
  select(
    age,
    sex,
    year_diagnosis,
    race,
    marriage_status,
    subtype,
    t_stage = tnm,
    n_stage,
    m_stage,
    grade,
    idc_type,
    radiation,
    chemotherapy,
    surgery,
    survival_months,
    survival_status,

  )

attach(seer)
```

```{r}

```

### 2024-11-05 final-project.Rmd
clean data

@@
   filter(
-    !is.na(subtype)
+    !is.na(subtype),
+    !is.na(grade),
+    n_stage != "NX"
   ) %>%

### 2024-11-07 final-project.Rmd
table 1

--- 2024-11-02 final-project.Rmd
@@
 library(psych)
+library(gmodels)
 ```
@@
 
+# there are soveral variables that need to be categorized for analysis and/filtered
+seer_cat <- seer_clean %>%
+   filter(
+    !is.na(subtype),
+    !is.na(grade),
+    n_stage != "NX"
+  ) %>%
+  mutate(agecat = case_when(
+    age %in% c("20-24 years", "25-29 years", "30-34 years", "35-39 years", "40-44 years", "45-49 years") ~ 0,
+    age %in% c("50-54 years", "55-59 years", "60-64 years", "65-69 years") ~ 1,
+    age %in% c("70-74 years", "75-79 years", "80-84 years", "85+ years") ~ 2,
+    TRUE ~ NA_real_  # Keeps other values as NA if they don’t match any of the above groups
+  ))
+
 # further filter and select only variables used in analysis
-seer <- seer_clean %>%
+seer <- seer_cat %>%
   filter(
-    !is.na(subtype)
+    !is.na(subtype),
+    !is.na(grade),
+    n_stage != "NX"
   ) %>%
   select(
-    age,
+    age = agecat,
     sex,
@@
     subtype,
-    t_stage = tnm,
     n_stage,
@@
     grade,
-    idc_type,
     radiation,
@@
 ```{r}
+# Table 1: Clinicopathological Characteristics by Breast Cancer Subtype
+# (unadjusted associations between covariates and exposure)
+
+## age
+CrossTable(age, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## race
+CrossTable(race, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## marriage status
+CrossTable(marriage_status, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## N stage
+CrossTable(n_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## M stage
+CrossTable(m_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## grade
+CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## radiation
+CrossTable(radiation, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## chemotherapy
+CrossTable(chemotherapy, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+
+## surgery
+CrossTable(surgery, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+```
+
+```{r}
+# Table 2: Clinicopathological Characteristics by survival
+# (Unadjusted associations between covariates and survival)
 

### 2024-11-07 final-project.Rmd
table 2 start

@@
 library(gmodels)
+library(survival)
 ```
@@
     grade2017 != "B-cell; pre-B; B-precursor",
-    grade2017 != "T-cell"
+    grade2017 != "T-cell",
+    idc_type %in% seq(8500, 8549)
   )
@@
 
-# there are soveral variables that need to be categorized for analysis and/filtered
+# there are several variables that need to be categorized for analysis and/filtered
 seer_cat <- seer_clean %>%
@@
     !is.na(subtype),
-    !is.na(grade),
-    n_stage != "NX"
+    #!is.na(grade),
+    #n_stage != "NX"
   ) %>%
@@
     TRUE ~ NA_real_  # Keeps other values as NA if they don’t match any of the above groups
+  )) %>%
+  mutate(gradecat = case_when(
+    grade %in% c(1, 2) ~ "I/II",
+    grade %in% c(3, 4) ~ "III/IV",
+    TRUE ~ NA_character_
+    )) %>%
+  # the survival_status column is what we use for event/censoring
+  # recode Dead as 1 and Alive (censored) as 0
+  mutate(status_cat = case_when(
+    survival_status == "Dead" ~ 1,
+    survival_status == "Alive" ~ 0
   ))
@@
 seer <- seer_cat %>%
-  filter(
-    !is.na(subtype),
-    !is.na(grade),
-    n_stage != "NX"
-  ) %>%
   select(
@@
     m_stage,
-    grade,
+    grade = gradecat,
     radiation,
@@
     survival_months,
-    survival_status,
+    survival_status = status_cat,
 
@@
 
+## age
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
 ```

### 2024-11-08 final-project.Rmd
table 2 and diagnostics

@@
 library(survival)
+library(survminer)
 ```
@@
       str_detect(n2015, "N3|N3N0S") ~ "N3",
-      str_detect(n2015, "NX") ~ "NX",
+      str_detect(n2015, "NX") ~ NA_character_,
       TRUE ~ NA_character_
@@
       str_detect(n2017, "c3|p3") ~ "N3",
-      str_detect(n2017, "cX") ~ "NX",
+      str_detect(n2017, "cX") ~ NA_character_,
       TRUE ~ NA_character_
@@
       str_detect(n2018, "N3|N3N0S") ~ "N3",
-      str_detect(n2018, "NX") ~ "NX",
+      str_detect(n2018, "NX") ~ NA_character_,
       TRUE ~ NA_character_
@@
 summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ age)
+
+## race
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ race), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ race)
+
+## marriage status
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ marriage_status)
+
+## N stage
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ n_stage)
+
+## M stage
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ m_stage)
+
+## grade
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ grade), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ grade)
+
+## radiation
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ radiation)
+
+## chemotherapy
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ chemotherapy)
+
+## surgery
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ surgery)
 ```
+
+```{r}
+# check the proportional hazards assumption for each covariate
+
+## age
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ age), data = seer, fun="cloglog")
+
+## race
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ race), data = seer, fun="cloglog")
+
+## marriage status
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), data = seer, fun="cloglog")
+
+## N stage
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), data = seer, fun="cloglog")
+
+## M stage
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), data = seer, fun="cloglog")
+
+## grade
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ grade), data = seer, fun="cloglog")
+
+## radiation
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), data = seer, fun="cloglog")
+
+## chemotherapy
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), data = seer, fun="cloglog")
+
+## surgery
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), data = seer, fun="cloglog")
+```

### 2024-11-13 final-project.Rmd
make categorical vars

@@
   )) %>%
+  mutate(agecat0 = case_when(
+    agecat == 0 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(agecat1 = case_when(
+    agecat == 1 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(agecat2 = case_when(
+    agecat == 2 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(racecat0 = case_when(
+    race == "White" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(racecat1 = case_when(
+    race == "Black" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(racecat2 = case_when(
+    race == "Other" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(marriage_status = case_when(
+    marriage_status == "Yes" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtypecat = case_when(
+    subtype == "Luminal A" ~ 0,
+    subtype == "Luminal B" ~ 1,
+    subtype == "Triple Negative" ~ 2,
+    subtype == "HER2 Positive" ~ 3,
+    TRUE ~ NA_integer_
+  )) %>%
+  mutate(subtype0 = case_when(
+    subtypecat == 0 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtype1 = case_when(
+    subtypecat == 1 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtype2 = case_when(
+    subtypecat == 2 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtype3 = case_when(
+    subtypecat == 3 ~ 1,
+    TRUE ~ 0
+  )) %>%
   mutate(gradecat = case_when(
-    grade %in% c(1, 2) ~ "I/II",
-    grade %in% c(3, 4) ~ "III/IV",
-    TRUE ~ NA_character_
+    grade %in% c(1, 2) ~ 0,
+    grade %in% c(3, 4) ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(n_stage = case_when(
+    n_stage %in% c("N0", "N1") ~ 0,
+    n_stage %in% c("N2", "N3") ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(m_stage = case_when(
+    m_stage == "M0" ~ 0,
+    m_stage == "M1" ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(radiation = case_when(
+    radiation == "No/Unknown" ~ 0,
+    radiation == "Yes" ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(chemotherapy = case_when(
+    chemotherapy == "No/Unknown" ~ 0,
+    chemotherapy == "Yes" ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(surgery = case_when(
+    surgery == "No/Unknown" ~ 0,
+    surgery == "Yes" ~ 1,
+    TRUE ~ NA_integer_
     )) %>%
@@
   select(
-    age = agecat,
+    agecat,
+    agecat0,
+    agecat1,
+    agecat2,
     sex,
@@
     race,
+    racecat0,
+    racecat1,
+    racecat2,
     marriage_status,
-    subtype,
+    subtype = subtypecat,
+    subtype0,
+    subtype1,
+    subtype2,
+    subtype3,
     n_stage,
@@
 
-## age
-CrossTable(age, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+## agecat
+CrossTable(agecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
 
@@
 
-## age
-summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
-survdiff(Surv(time = survival_months, event = survival_status) ~ age)
+## agecat
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ agecat)
 
@@
 ```{r}
-# check the proportional hazards assumption for each covariate
+# check the proportional hazards assumption for each covariate with (1) a log log plot
 
-## age
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ age), data = seer, fun="cloglog")
+## agecat
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), data = seer, fun="cloglog")
 
@@
 ```
+
+```{r}
+# create a dataset in counting process format
+seer.cp <- survSplit(seer, cut=seer$survival_months[seer$survival_status==1],
+                     end="survival_months",event="survival_status", start="start")
+
+```
+
+```{r}
+# check PH assumption for race, n_stage, chemotherapy, surgery with statistical approach
+
+# create time-varying covariates for race by adding adding a product term with the ln of time
+seer.cp$logtrace1=seer.cp$race[seer.cp$race == "White"]*log(seer.cp$survival_months)
+
+seer.cp$logtrace <- ifelse(seer.cp$race == "White", log(seer.cp$survival_months), NA)
+
+```
+
+```{r}
+model <- coxph(
+  Surv(start, survival_months, survival_status) ~ subtype0 + subtype1 + subtype2 + subtype3 +  
+    agecat0 + agecat1 + agecat2 + 
+    racecat0 + racecat1 + racecat2 +
+    marriage_status +
+
+               factor(SEX) + factor(cholcat) + CURSMOKE + bmicat1 + bmicat2, data = fham.cp3)
+```

### 2024-11-13 final-project.Rmd
ph assumption

--- 2024-11-08 final-project.Rmd
@@
 )
+```
 
+```{r}
 # filter
@@
   )
+```
 
+```{r}
 # clean up the variables
@@
     )
+```
 
+```{r}
 # there are several variables that need to be categorized for analysis and/filtered
@@
   )) %>%
+  mutate(agecat0 = case_when(
+    agecat == 0 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(agecat1 = case_when(
+    agecat == 1 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(agecat2 = case_when(
+    agecat == 2 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(racecat = case_when(
+    race == "White" ~ 0,
+    race == "Black" ~ 1,
+    race == "Other" ~ 2,
+    TRUE ~ NA_integer_
+  )) %>%
+  mutate(racecat0 = case_when(
+    race == "White" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(racecat1 = case_when(
+    race == "Black" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(racecat2 = case_when(
+    race == "Other" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(marriage_status = case_when(
+    marriage_status == "Yes" ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtypecat = case_when(
+    subtype == "Luminal A" ~ 0,
+    subtype == "Luminal B" ~ 1,
+    subtype == "Triple Negative" ~ 2,
+    subtype == "HER2 Positive" ~ 3,
+    TRUE ~ NA_integer_
+  )) %>%
+  mutate(subtype0 = case_when(
+    subtypecat == 0 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtype1 = case_when(
+    subtypecat == 1 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtype2 = case_when(
+    subtypecat == 2 ~ 1,
+    TRUE ~ 0
+  )) %>%
+  mutate(subtype3 = case_when(
+    subtypecat == 3 ~ 1,
+    TRUE ~ 0
+  )) %>%
   mutate(gradecat = case_when(
-    grade %in% c(1, 2) ~ "I/II",
-    grade %in% c(3, 4) ~ "III/IV",
-    TRUE ~ NA_character_
+    grade %in% c(1, 2) ~ 0,
+    grade %in% c(3, 4) ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(n_stage = case_when(
+    n_stage %in% c("N0", "N1") ~ 0,
+    n_stage %in% c("N2", "N3") ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(m_stage = case_when(
+    m_stage == "M0" ~ 0,
+    m_stage == "M1" ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(radiation = case_when(
+    radiation == "No/Unknown" ~ 0,
+    radiation == "Yes" ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(chemotherapy = case_when(
+    chemotherapy == "No/Unknown" ~ 0,
+    chemotherapy == "Yes" ~ 1,
+    TRUE ~ NA_integer_
+    )) %>%
+  mutate(surgery = case_when(
+    surgery == "No/Unknown" ~ 0,
+    surgery == "Yes" ~ 1,
+    TRUE ~ NA_integer_
     )) %>%
@@
   ))
+```
 
+```{r}
 # further filter and select only variables used in analysis
@@
   select(
-    age = agecat,
-    sex,
-    year_diagnosis,
-    race,
+    agecat,
+    agecat0,
+    agecat1,
+    agecat2,
+    racecat,
+    racecat0,
+    racecat1,
+    racecat2,
     marriage_status,
-    subtype,
+    subtype = subtypecat,
+    subtype0,
+    subtype1,
+    subtype2,
+    subtype3,
     n_stage,
@@
 ```{r}
-# check the proportional hazards assumption for each covariate
+# check the proportional hazards assumption for each covariate with (1) a log log plot
 
-## age
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ age), data = seer, fun="cloglog")
+## subtype
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ subtype), data = seer, fun="cloglog")
+
+## agecat
+ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), data = seer, fun="cloglog")
 
@@
 ```
+
+```{r}
+# create a dataset in counting process format
+seer.cpfull <- survSplit(seer, cut=seer$survival_months[seer$survival_status==1],
+                     end="survival_months",event="survival_status", start="start")
+# remove NA because the model will do this anyways
+seer.cp <- na.omit(seer.cpfull)
+```
+
+```{r}
+# check PH assumption for race, n_stage, chemotherapy, surgery with statistical approach
+
+# create time-varying covariates for race by adding adding a product term with the ln of time
+seer.cp$logtrace1=seer.cp$race[seer.cp$race == "White"]*log(seer.cp$survival_months)
+
+seer.cp$logtrace <- ifelse(seer.cp$race == "White", log(seer.cp$survival_months), NA)
+
+```
+
+```{r}
+model <- coxph(
+  Surv(start, survival_months, survival_status) ~ relevel(factor(subtype), ref = '0') +
+    relevel(factor(agecat), ref = '0') + 
+    relevel(factor(racecat), ref = '0') +
+    relevel(factor(marriage_status),ref='1') + 
+    relevel(factor(n_stage),ref='0') + 
+    relevel(factor(m_stage),ref='0') +
+    relevel(factor(grade), ref='0') +
+    relevel(factor(radiation), ref='1') + 
+    relevel(factor(chemotherapy), ref='1') +
+    relevel(factor(surgery), ref='1'),
+  data = seer.cp, 
+  method = "breslow")
+
+summary(model)
+
+# Test proportional hazards assumption with Schoenfeld residuals
+cox.zph(model)
+```

### 2024-11-22 code/final-project.Rmd
progress

---
title: "final-project"
author: "Murphy John"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(stringr)
library(psych)
library(gmodels)
library(survival)
library(survminer)
```

```{r}
# load data
seer <- readRDS("data/processed/seer.rds")
attach(seer)
```

```{r}
# Table 1: Clinicopathological Characteristics by Breast Cancer Subtype
# (unadjusted associations between covariates and exposure)

## agecat
CrossTable(agecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## race
CrossTable(racecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## marriage status
CrossTable(marriage_status, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## N stage
CrossTable(n_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## M stage
CrossTable(m_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## grade
CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE, useNA = "always")

## radiation
CrossTable(radiation, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## chemotherapy
CrossTable(chemotherapy, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## surgery
CrossTable(surgery, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

```{r}
# Table 2: Clinicopathological Characteristics by survival
# (Unadjusted associations between covariates and survival) also median survival

## agecat
survfit(Surv(survival_months, survival_status)~agecat)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ agecat)

## race
survfit(Surv(survival_months, survival_status)~racecat)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ racecat), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ racecat)

## marriage status
survfit(Surv(survival_months, survival_status)~marriage_status)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ marriage_status)

## N stage
survfit(Surv(survival_months, survival_status)~n_stage)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ n_stage)

## M stage
survfit(Surv(survival_months, survival_status)~m_stage)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ m_stage)

## grade
survfit(Surv(survival_months, survival_status)~grade)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ grade), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ grade)

## radiation
survfit(Surv(survival_months, survival_status)~radiation)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ radiation)

## chemotherapy
survfit(Surv(survival_months, survival_status)~chemotherapy)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ chemotherapy)

## surgery
survfit(Surv(survival_months, survival_status)~surgery)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ surgery)
```

```{r}
# Figure 2: Log-log plots to check Proportional Hazards assumption
library(gridExtra)
library(survminer)
library(survival)

plot_subtype <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ subtype), 
  data = seer, 
  fun ="cloglog",
  ylab = "Log-log survival probability",
  xlab = "Time (months)",
  legend.labs = c("Luminal A", "Luminal B", "HER2 Positive", "Triple Negative"))

plot_agecat <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ agecat), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("20-49 years", "50-69 years", ">= 70 years"))

plot_race <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ racecat), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("White", "Black", "Other"))

plot_marriage <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("Unmarried", "Married"))

plot_n_stage <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("N0/1", "N2/3"))

plot_m_stage <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("M0", "M1"))

plot_grade <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ grade), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("I/II", "III/IV"))

plot_radiation <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ radiation), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("No/Unknown", "Yes"))

plot_chemotherapy <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("No/Unknown", "Yes"))

plot_surgery <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ surgery), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("No/Unknown", "Yes"))

grid.arrange(plot_subtype$plot, plot_agecat$plot, plot_race$plot, plot_marriage$plot, plot_n_stage$plot,
             plot_m_stage$plot, plot_grade$plot, plot_radiation$plot, plot_chemotherapy$plot, plot_surgery$plot,
             ncol = 5, nrow = 2)
```

```{r}
# check the proportional hazards assumption for each covariate with (1) a log log plot

## subtype
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ subtype), data = seer, fun="cloglog")

## agecat
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), data = seer, fun="cloglog")

## race
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ racecat), data = seer, fun="cloglog")

## marriage status
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), data = seer, fun="cloglog")

## N stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), data = seer, fun="cloglog")

## M stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), data = seer, fun="cloglog")

## grade
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ grade), data = seer, fun="cloglog")

## radiation
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), data = seer, fun="cloglog")

## chemotherapy
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), data = seer, fun="cloglog")

## surgery
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), data = seer, fun="cloglog")
```

```{r}
# create a dataset in counting process format
seer.cpfull <- survSplit(seer, cut=seer$survival_months[seer$survival_status==1],
                     end="survival_months",event="survival_status", start="start")
# remove NA because the model will do this anyways
seer.cp <- na.omit(seer.cpfull)
```

```{r}
model <- coxph(
  Surv(survival_months, survival_status) ~ 
    relevel(factor(subtype), ref = '0') +
    relevel(factor(agecat), ref = '0') + 
    relevel(factor(racecat), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1'),
  data = seer, 
  method = "breslow")

summary(model)

# Test proportional hazards assumption with Schoenfeld residuals
resid <- cox.zph(model, terms = FALSE)

# get the correlations between residuals and ranked survival times
apply(resid$y, 2, \(x) cor.test(rank(resid$time), x))
# subtype1: -0.034
# subtype2: -0.1296
# subtype3: -0.0435
# agecat1: -0.0331
# agecat2: -0.0304
# racecat1: 0.0077
# racecat2: 0.0102
# marraige_status0: 0.0043
# n_stage1: 0.0188
# m_stage: -0.0046
# grade: -0.0412
# radiation: -0.072
# chemotherapy: -0.1122
# surgery: -0.0715

cor <- apply(resid$y, 2, \(x) {
  test <- cor.test(rank(resid$time), x)  # Perform correlation test
  c(
    correlation = test$estimate,
    p_value = test$p.value,
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2]
  ) 
})

cor_df <- as.data.frame(t(cor))
colnames(cor_df) <- c("Correlation", "P_Value", "Conf_Low", "Conf_High")
rownames(cor_df) <- c(
  "subtype1",
  "subtype2",
  "subtype3",
  "agecat1",
  "agecat2",
  "racecat1",
  "racecat2",
  "marraige_status0",
  "n_stage1",
  "m_stage",
  "grade",
  "radiation",
  "chemotherapy",
  "surgery"
)

```

```{r}
# fit extended model and see if interaction terms are significant
# create a dataset in counting process format
seer.cpfull <- survSplit(seer, cut=seer$survival_months[seer$survival_status==1],
                     end="survival_months",event="survival_status", start="start")
# remove NA because the model will do this anyways
seer.cp <- na.omit(seer.cpfull)

# create interaction between each subtype and log time
seer.cp$subtype1_time=seer.cp$subtype1*log(seer.cp$survival_months)
seer.cp$subtype2_time=seer.cp$subtype2*log(seer.cp$survival_months)
seer.cp$subtype3_time=seer.cp$subtype3*log(seer.cp$survival_months)

options(scipen=999)
model_ext <- coxph(
  formula = Surv(
    seer.cp$start, seer.cp$survival_months, seer.cp$survival_status) ~ 
    subtype1 + subtype2 + subtype3 + 
    relevel(factor(agecat), ref = '0') + 
    relevel(factor(racecat), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1') +
    subtype1_time + subtype2_time + subtype3_time, 
  data = seer.cp, method="breslow")

summary(model_ext)

```

### 2024-11-22 code/analysis/flowchart.R

@@ -1,0 +1,63 @@
+# Figure 1: Flowchart of IBC patients screening on SEER database
+
+library(Gmisc, quietly = TRUE)
+library(glue)
+library(htmlTable)
+library(grid)
+library(magrittr)
+
+
+eligible <- boxGrob(glue("Breast cencer patients diagnosed with \ninflammatory breast carcinoma between 2010-2021",
+                         "N = {pop}",
+                         pop = txtInt(5986),
+                         .sep = "\n"))
+
+included <- boxGrob(glue("Eligible inflammatory breast \ncancer patients",
+                         "n = {incl}",
+                         incl = txtInt(4140),
+                         .sep = "\n"))
+
+excluded <- boxGrob(glue("Excluded (n = {tot}):",
+                         " - Age at diagnosis < 20 years: {age}",
+                         " - Male sex: {male}",
+                         " - Survival time < 1 month: {survival}",
+                         " - Unknown or competing \ncause of death: {cod}",
+                         " - Incomplete or unknown \nclinicopathological characteristics: {incomplete}",
+                         tot = txtInt(1846),
+                         age = 1,
+                         male = 16,
+                         survival = 165,
+                         cod = 487,
+                         incomplete = txtInt(1177),
+                         .sep = "\n"),
+                    just = "left")
+
+png("results/figures/flowchart.png", width = 700, height = 500)
+
+grid.newpage()
+
+vert <- spreadVertical(eligible = eligible,
+                       included = included,
+                       .from = .9,
+                       .to = .1,
+                       .type = "center")
+
+excluded <- moveBox(excluded,
+                    x = .8,
+                    y = coords(vert$included)$top + 
+                      distance(vert$eligible, 
+                               vert$included, half = TRUE, center = FALSE))
+
+for (i in 1:(length(vert) - 1)) {
+  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
+    print
+}
+
+connectGrob(vert$eligible, excluded, type = "L")
+grid.rect(gp = gpar(col = "black", fill = NA, lwd = 2))  # Adjust color and line width as needed
+
+# Print boxes
+vert
+excluded
+
+dev.off()

### 2024-11-22 code/processing/processing-script.R

@@ -1,0 +1,325 @@
+
+library(dplyr)
+library(stringr)
+library(psych)
+library(gmodels)
+library(survival)
+library(survminer)
+
+# load data
+seer1 <- read.csv("data/raw/caselist_2024.11.csv", header = TRUE)
+
+# column rename
+seer_name <- seer1 %>%
+  dplyr::rename(
+    age = Age.recode.with..1.year.olds,
+    sex = Sex,
+    year_diagnosis = Year.of.diagnosis,
+    race = Race.ethnicity,
+    marriage_status = Marital.status.at.diagnosis,
+    tnm2015 = Derived.AJCC.T..7th.ed..2010.2015.,
+    tnm2017 = Derived.SEER.Combined.T..2016.2017.,
+    tnm2018 = Derived.EOD.2018.T..2018..,
+    subtype = Breast.Subtype..2010..,
+    grade2017 = Grade.Recode..thru.2017.,
+    grade2018 = Derived.Summary.Grade.2018..2018..,
+    n2015 = Derived.AJCC.N..7th.ed..2010.2015.,
+    n2017 = Derived.SEER.Combined.N..2016.2017.,
+    n2018 = Derived.EOD.2018.N..2018..,
+    m2015 = Derived.AJCC.M..7th.ed..2010.2015.,
+    m2017 = Derived.SEER.Combined.M..2016.2017.,
+    m2018 = Derived.EOD.2018.M..2018..,
+    radiation = Radiation.recode,
+    chemotherapy = Chemotherapy.recode..yes..no.unk.,
+    surgery = RX.Summ..Surg.Prim.Site..1998..,
+    survival_months = Survival.months,
+    survival_status = Vital.status.recode..study.cutoff.used.,
+    survival_cause_cancer = SEER.cause.specific.death.classification,
+    survival_cause_other = SEER.other.cause.of.death.classification,
+    idc_type = Histologic.Type.ICD.O.3
+  )
+
+seer_idc <- seer_name %>%
+  filter(
+    idc_type %in% seq(8500, 8549) # 481 indiv exclude, 5986 remain
+  )
+
+# filter by exclusion criteria
+seer_filter <- seer_idc %>%
+  filter(
+    age != "00 years",
+    age != "15-19 years", # 1 indiv exclude, 5985 remain
+    sex == "Female", # 16 indiv exclude, 5969 remain
[…]

### 2024-11-22 code/final-project.Rmd
table 3 updates

@@
 ```{r}
-# Figure 2: Log-log plots to check Proportional Hazards assumption
-library(gridExtra)
-library(survminer)
-library(survival)
-
-plot_subtype <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ subtype), 
-  data = seer, 
-  fun ="cloglog",
-  ylab = "Log-log survival probability",
-  xlab = "Time (months)",
-  legend.labs = c("Luminal A", "Luminal B", "HER2 Positive", "Triple Negative"))
-
-plot_agecat <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ agecat), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("20-49 years", "50-69 years", ">= 70 years"))
-
-plot_race <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ racecat), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("White", "Black", "Other"))
-
-plot_marriage <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("Unmarried", "Married"))
-
-plot_n_stage <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("N0/1", "N2/3"))
-
-plot_m_stage <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("M0", "M1"))
-
-plot_grade <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ grade), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("I/II", "III/IV"))
-
-plot_radiation <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ radiation), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("No/Unknown", "Yes"))
-
-plot_chemotherapy <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("No/Unknown", "Yes"))
-
-plot_surgery <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ surgery), 
-  data = seer, 
-  fun="cloglog",
-  ylab= "Log-log survival probability",
-  xlab= "Time (months)",
-  legend.labs = c("No/Unknown", "Yes"))
-
-grid.arrange(plot_subtype$plot, plot_agecat$plot, plot_race$plot, plot_marriage$plot, plot_n_stage$plot,
-             plot_m_stage$plot, plot_grade$plot, plot_radiation$plot, plot_chemotherapy$plot, plot_surgery$plot,
-             ncol = 5, nrow = 2)
-```
-
-```{r}
-# check the proportional hazards assumption for each covariate with (1) a log log plot
-
-## subtype
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ subtype), data = seer, fun="cloglog")
-
-## agecat
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), data = seer, fun="cloglog")
-
-## race
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ racecat), data = seer, fun="cloglog")
-
-## marriage status
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), data = seer, fun="cloglog")
-
-## N stage
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), data = seer, fun="cloglog")
-
-## M stage
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), data = seer, fun="cloglog")
-
-## grade
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ grade), data = seer, fun="cloglog")
-
-## radiation
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), data = seer, fun="cloglog")
-
-## chemotherapy
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), data = seer, fun="cloglog")
-
-## surgery
-ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), data = seer, fun="cloglog")
-```
-
-```{r}
 # create a dataset in counting process format
@@
 ```{r}
+# check the correlation between schoenfeld residuals and ranked survival times
+
+# run the cox ph model
 model <- coxph(
@@
 summary(model)
+
+options(scipen=999)
 
@@
   c(
-    correlation = test$estimate,
-    p_value = test$p.value,
-    conf_low = test$conf.int[1],
-    conf_high = test$conf.int[2]
+    correlation = round(test$estimate, 3),
+    p_value = round(test$p.value, 3),
+    conf_low = round(test$conf.int[1], 3),
+    conf_high = round(test$conf.int[2], 3)
   ) 
@@
 
-cor_df <- as.data.frame(t(cor))
-colnames(cor_df) <- c("Correlation", "P_Value", "Conf_Low", "Conf_High")
+# Convert the matrix to a data frame
+cor_df_all <- as.data.frame(t(cor))
+
+# Now, we can correctly mutate and format the output
+cor_df <- cor_df_all %>%
+  mutate(
+    Correlation = paste(correlation.cor, " (", conf_low, ", ", conf_high, ")", sep = ""),
+    P_Value = as.character(p_value)
+  ) %>%
+  select(Correlation, P_Value)
 rownames(cor_df) <- c(
@@
 )
+
+# table 3: Correlation between Schoenfeld residuals and ranked survival time
+t3_raw <- cor_df %>%
+  add_row(Correlation = "", P_Value = "", .before = 1) %>%
+  add_row(Correlation = "", P_Value = "", .before = 1) %>%
+  add_row(Correlation = "", P_Value = "", .before = 6) %>%
+  add_row(Correlation = "", P_Value = "", .before = 6) %>%
+  add_row(Correlation = "", P_Value = "", .before = 10) %>%
+  add_row(Correlation = "", P_Value = "", .before = 10) %>%
+  add_row(Correlation = "", P_Value = "", .before = 14) %>%
+  add_row(Correlation = "", P_Value = "", .before = 14) %>%
+  add_row(Correlation = "", P_Value = "", .before = 17) %>%
+  add_row(Correlation = "", P_Value = "", .before = 17) %>%
+  add_row(Correlation = "", P_Value = "", .before = 20) %>%
+  add_row(Correlation = "", P_Value = "", .before = 20) %>%
+  add_row(Correlation = "", P_Value = "", .before = 23) %>%
+  add_row(Correlation = "", P_Value = "", .before = 23) %>%
+  add_row(Correlation = "", P_Value = "", .before = 26) %>%
+  add_row(Correlation = "", P_Value = "", .before = 26) %>%
+  add_row(Correlation = "", P_Value = "", .before = 29) %>%
+  add_row(Correlation = "", P_Value = "", .before = 29) %>%
+  add_row(Correlation = "", P_Value = "", .before = 32) %>%
+  add_row(Correlation = "", P_Value = "", .before = 32)
 

### 2024-11-22 code/analysis/loglog-plots.R

@@ -1,0 +1,110 @@
+# Figure 2: Log(-log) plots to check Proportional Hazards assumption
+
+library(gridExtra)
+library(survminer)
+library(survival)
+
+# load data
+seer <- readRDS("data/processed/seer.rds")
+attach(seer)
+
+# create each plot
+plot_subtype <- ggsurvplot(
+  survfit(Surv(time = survival_months, event = survival_status) ~ subtype), 
+  data = seer, 
+  fun ="cloglog",
+  ylab = "log(-log(S(t))",
+  xlab = "log(Time)",
+  legend.labs = c("Luminal A", "Luminal B", "HER2 Positive", "Triple Negative"),
+  legend.title = "Subtype:")
+
+plot_agecat <- ggsurvplot(
+  survfit(Surv(time = survival_months, event = survival_status) ~ agecat), 
+  data = seer, 
+  fun="cloglog",
+  ylab = "log(-log(S(t))",
+  xlab = "log(Time)",
+  legend.labs = c("20-49 years", "50-69 years", ">= 70 years"),
+  legend.title = "Age:")
+
+plot_race <- ggsurvplot(
+  survfit(Surv(time = survival_months, event = survival_status) ~ racecat), 
+  data = seer, 
+  fun="cloglog",
+  ylab = "log(-log(S(t))",
+  xlab = "log(Time)",
+  legend.labs = c("White", "Black", "Other"),
+  legend.title = "Race:")
+
+plot_marriage <- ggsurvplot(
+  survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), 
+  data = seer, 
+  fun="cloglog",
+  ylab = "log(-log(S(t))",
+  xlab = "log(Time)",
+  legend.labs = c("Unmarried", "Married"),
+  legend.title = "Marital Status:")
+
+plot_n_stage <- ggsurvplot(
+  survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), 
+  data = seer, 
+  fun="cloglog",
+  ylab = "log(-log(S(t))",
+  xlab = "log(Time)",
+  legend.labs = c("N0/1", "N2/3"),
+  legend.title = "N Stage:")
+
+plot_m_stage <- ggsurvplot(
+  survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), 
+  data = seer, 
+  fun="cloglog",
+  ylab = "log(-log(S(t))",
+  xlab = "log(Time)",
[…]

### 2024-11-22 code/final-project.Rmd
table 1

@@
 # (Unadjusted associations between covariates and survival) also median survival
+# this is univariate cox ph model
+# Risk factors for OS according to the Cox proportional hazards regression model.
 
@@
 # check the correlation between schoenfeld residuals and ranked survival times
+library(survival)
+library(dplyr)
 
 # run the cox ph model
+# this will be in Table 2 as multivariate analysis
 model <- coxph(

### 2024-11-22 code/analysis/table1.R

@@ -1,0 +1,73 @@
+# Table 1: Clinicopathological Characteristics by Breast Cancer Subtype
+# (unadjusted associations between covariates and exposure)
+
+# Declare dependencies
+box::use(
+  readr,
+  here,
+  gt,
+  gtExtras,
+  gtsummary,
+  modelsummary,
+  dplyr,
+  webshot2,
+  tidyr,
+  ggplot2,
+  Hmisc
+)
+
+# load data
+seer <- readRDS("data/processed/seer_tab.rds")
+attach(seer)
+
+t1 <-
+  seer |>
+  gtsummary::tbl_summary(
+    by = subtype,
+    include = c(agecat, race, marriage_status, n_stage, m_stage,
+                gradecat, radiation, chemotherapy, surgery),
+    label = list(
+      agecat ~ "Age (years)",
+      race ~ "Race",
+      marriage_status ~ "Marital Status",
+      n_stage ~ "N Stage",
+      m_stage ~ "M Stage",
+      gradecat ~ "Grade",
+      radiation ~ "Radiation",
+      chemotherapy ~ "Chemotherapy",
+      surgery ~ "Surgery"
+    ),
+    type = list(
+      marriage_status ~ "categorical",
+      n_stage ~ "categorical",
+      m_stage ~ "categorical",
+      gradecat ~ "categorical",
+      radiation ~ "categorical",
+      chemotherapy ~ "categorical",
+      surgery ~ "categorical"
+    ),
+    digits = c(agecat, race, marriage_status, n_stage, m_stage,
+               gradecat, radiation, chemotherapy, surgery) ~ c(0, 1),
+    statistic_header = "**Statistic**"
+  ) |>
+  gtsummary::add_overall() |>
+  gtsummary::add_p() |>
+  gtsummary::modify_caption(
+    "Table 1: Clinicopathological characteristics by cancer subtype"
+  ) |>
+  gtsummary::as_gt()
+
+gt::gtsave(
+  data = t1,
+  filename = here::here("results/tables", "t1.png"),
+  vwidth = 2400,
+  vheight = 1350
+)
+
+t4b |>
+  gt::rm_caption() |>
+  gt::gtsave(
+    filename = here::here("results", "paper-t1.png"),
+    vwidth = 2400,
+    vheight = 1350
+  )

### 2024-11-22 code/processing/processing-script.R

@@ -190,6 +190,61 @@ seer_complete <- seer_clean %>%
     n_stage != "NX" # 106 indiv excluded, 4140 remain
   )
 
+# there are several variables that need to be categorized for analysis and/filtered
+# for this df leave the category names legible for creating tables
+seer_catclean <- seer_complete %>%
+  mutate(subtype = factor(
+    subtype, levels = c("Luminal A", "Luminal B", "HER2 Positive", "Triple Negative")
+  )) %>%
+  mutate(agecat = case_when(
+    age %in% c("20-24 years", "25-29 years", "30-34 years", "35-39 years", "40-44 years", "45-49 years") ~ "20-49 years",
+    age %in% c("50-54 years", "55-59 years", "60-64 years", "65-69 years") ~ "50-69 years",
+    age %in% c("70-74 years", "75-79 years", "80-84 years", "85+ years") ~ ">=70 years",
+    TRUE ~ NA_character_  # Keeps other values as NA if they don’t match any of the above groups
+  )) %>%
+  mutate(agecat = factor(
+    agecat, levels = c("20-49 years", "50-69 years", ">=70 years")
+  )) %>%
+  mutate(race = factor(
+      race, levels = c("White", "Black", "Other")
+  )) %>%
+  mutate(marriage_status = case_when(
+    marriage_status == "Yes" ~ "Married",
+    TRUE ~ "Unmarried"
+  )) %>%
+  mutate(marriage_status = factor(
+    marriage_status, levels = c("Unmarried", "Married")
+  )) %>%
+  mutate(gradecat = case_when(
+    grade %in% c(1, 2) ~ "I/II",
+    grade %in% c(3, 4) ~ "III/IV",
+    TRUE ~ NA_character_
+  )) %>%
+  mutate(n_stage = case_when(
+    n_stage %in% c("N0", "N1") ~ "N0/N1",
+    n_stage %in% c("N2", "N3") ~ "N2/N3",
+    TRUE ~ NA_character_
+  ))
+
+# save this data with named categories as it will be useful for creating tables
+seer_tab <- seer_catclean %>%
+  select(
+    agecat,
+    race,
+    marriage_status,
+    subtype,
+    n_stage,
+    m_stage,
+    gradecat,
+    radiation,
+    chemotherapy,
+    surgery,
+    survival_months,
+    survival_status,
+  )
+
+# save as rds file
+saveRDS(seer_tab, file = "data/processed/seer_tab.rds")
 
[…]

### 2024-11-25 code/final-project.Rmd
tables

@@
 
-## agecat
-CrossTable(agecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+## age
+CrossTable(age, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
 
 ## race
-CrossTable(racecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
+CrossTable(race, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
 
@@
 ## grade
-CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE, useNA = "always")
+CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
 
@@
 
-## agecat
-survfit(Surv(survival_months, survival_status)~agecat)
-summary(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), times = c(12, 60, 120))
-survdiff(Surv(time = survival_months, event = survival_status) ~ agecat)
+## age
+survfit(Surv(survival_months, survival_status)~age)
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ age)
 
 ## race
-survfit(Surv(survival_months, survival_status)~racecat)
-summary(survfit(Surv(time = survival_months, event = survival_status) ~ racecat), times = c(12, 60, 120))
-survdiff(Surv(time = survival_months, event = survival_status) ~ racecat)
+survfit(Surv(survival_months, survival_status)~race)
+summary(survfit(Surv(time = survival_months, event = survival_status) ~ race), times = c(12, 60, 120))
+survdiff(Surv(time = survival_months, event = survival_status) ~ race)
 
@@
     relevel(factor(subtype), ref = '0') +
-    relevel(factor(agecat), ref = '0') + 
-    relevel(factor(racecat), ref = '0') +
+    relevel(factor(age), ref = '0') + 
+    relevel(factor(race), ref = '0') +
     relevel(factor(marriage_status),ref='1') + 
@@
 # subtype3: -0.0435
-# agecat1: -0.0331
-# agecat2: -0.0304
-# racecat1: 0.0077
-# racecat2: 0.0102
+# age1: -0.0331
+# age2: -0.0304
+# race1: 0.0077
+# race2: 0.0102
 # marraige_status0: 0.0043
@@
   "subtype3",
-  "agecat1",
-  "agecat2",
-  "racecat1",
-  "racecat2",
+  "age1",
+  "age2",
+  "race1",
+  "race2",
   "marraige_status0",
   "n_stage1",
+  "n_stage2",
+  "n_stage3",
   "m_stage",
@@
     subtype1 + subtype2 + subtype3 + 
-    relevel(factor(agecat), ref = '0') + 
-    relevel(factor(racecat), ref = '0') +
+    relevel(factor(age), ref = '0') + 
+    relevel(factor(race), ref = '0') +
     relevel(factor(marriage_status),ref='1') + 

### 2024-11-25 products/manuscript/manuscript.qmd
tables

---
title: "epid 8010 manuscript"
author: " "
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	error = FALSE,
	knitr.graphics.error = FALSE
)
library(yaml, include.only = NULL)
library(knitr, include.only = NULL)
library(here)

# Loading data to present in text
seer <- readr::read_rds(here::here("data", "processed", "seer.rds"))
```

**Authors**

-   Murphy John$^{1}$

**Author affiliation**

1.  Department of Epidemiology and Biostatistics, University of Georgia, Athens, GA, USA.

\pagebreak

# Abstract

**Background:**

**Methods:**

**Results:**

**Conclusions:**

**Keywords:**

\pagebreak

# Introduction

Inflammatory breast carcinoma (IBC) is a rare but aggressive subtype of breast cancer. Reportedly, IBC accounts for only 1% to 6% of breast cancer cases in the United States, yet it was responsible for 10% of all types of breast cancer-related deaths (Zhu)

Inflammatory breast cancer (IBC) is a rare and aggressive variant of stage IIID breast cancer, with increased likelihood of metastasis upon diagnosis relative to non-inflammatory breast cancer. (Abraham)

Inflammatory breast cancer (IBC) is undoubtedly a unique biologic subtype of breast cancer with poor prognosis, which accounts for 1e5% of all breast cancer and comprises 7.0% of all breast cancerspecific mortality. (Pan)

According to the TNM classification, IBC is classified as T4d (a category reserved for IBC) for its aggressive biological behaviour. With great efforts to improve survival, including advancements in surgical technique, targeted therapy, radiotherapy and chemotherapy and improvements in supportive treatment, the prognosis of IBC has increased (Pan)

Although IBC demonstrates an inflammatory appearance, it is caused by cancer cells obstructing the dermal lymphatics of the breasts. (Zhu)

IBC is categorized as T4d (a category designated for IBC) in the TNM classification with its biologically aggressive character. With the application of a multimodal therapy strategy that includes systemic neoadjuvant chemotherapy, surgery, and radiotherapy, the clinical outcomes of IBC have significantly improved. (Zhu)

However, the approximately 35% to 40% 5-year cumulative survival rate is much lower compared with other breast cancers. (Zhu)

Historically, single-modality treatment to cure IBC was not successful; 90% of patients developed recurrent and/or metastatic disease within 2 years, and the 5-year survival rate was 5%. Combinations of neoadjuvant systemic chemotherapy, surgery, and radiation therapy have led to an improved prognosis. However, the overall 5-year survival rate for patients with IBC is still very low, at 30% (Yamauchi)

The aim of this paper is to determine the effect of cancer subtype on survival in adult femaes with IBC.

"T4 (includes T4a, T4b, T4c, and T4d): Tumor of any size growing into the chest wall or skin. This includes inflammatory breast cancer." https://www.cancer.org/cancer/types/breast-cancer/understanding-a-breast-cancer-diagnosis/stages-of-breast-cancer.html#:~:text=T1%20(includes%20T1a%2C%20T1b%2C,the%20chest%20wall%20or%20skin.

"Inflammatory breast cancer (IBC) is rare. It accounts for only 1% to 5% of all breast cancers. Although it is a type of invasive ductal carcinoma, its symptoms, outlook, and treatment are different. IBC causes symptoms of breast inflammation like swelling and redness, which is caused by cancer cells blocking lymph vessels in the skin causing the breast to look "inflamed." " https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html

"IBC doesn't look like a typical breast cancer. It often does not cause a breast lump, and it might not show up on a mammogram. This makes it harder to diagnose., IBC tends to occur in younger women (younger than 40 years of age)., Black women appear to develop IBC more often than White women., IBC is more common among women who are overweight or obese. IBC tends to be more aggressive—it grows and spreads much more quickly—than more common types of breast cancer., IBC is always at least at a locally advanced stage when it’s first diagnosed because the breast cancer cells have grown into the skin. (This means it is at least stage III.), In about 1 of 3 cases, IBC has already spread (metastasized) to distant parts of the body when it is diagnosed. This makes it harder to treat successfully., Women with IBC tend to have a worse prognosis (outcome) than women with other common types of breast cancer." https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html

"All inflammatory breast cancers start as stage III (T4dNXM0) since they involve the skin. If the cancer has spread outside the breast to distant parts of the body, it is stage IV." https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html

"Inflammatory breast cancer (IBC) that has not spread outside the breast is stage III. In most cases, treatment is chemotherapy first to try to shrink the tumor, followed by surgery to remove the cancer. Radiation and often other treatments, like more chemotherapy or targeted drug therapy, are given after surgery. Because IBC is so aggressive, breast conserving surgery (lumpectomy) and sentinel lymph node biopsy are typically not part of the treatment. IBC that has spread to other parts of the body (stage IV) may be treated with chemotherapy, hormone therapy, and/or targeted drugs." https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html

"The most common and widely accepted classification of breast cancer is from an immunohistochemical perspective, based on the expression of the following hormone receptors: estrogen (ER), progesterone (PR) and human epidermal growth factor (HER2). Accordingly, the following four subtypes of breast cancer are widely recognized: luminal A, luminal B, HER2-positive, and triple-negative. " https://www.ncbi.nlm.nih.gov/books/NBK583808/#:~:text=Accordingly%2C%20the%20following%20four%20subtypes,positive%2C%20and%20triple%2Dnegative.

"Breast cancer has four primary molecular subtypes, defined in large part by hormone receptors (HR) and other types of proteins involved (or not involved) in each cancer: Luminal A or HR+/HER2- (HR-positive/HER2-negative) Luminal B or HR+/HER2+ (HR-positive/HER2-positive) Triple-negative or HR-/HER2- (HR/HER2-negative) HER2-positive" https://www.cancercenter.com/cancer-types/breast-cancer/types/breast-cancer-molecular-types#:~:text=Breast%20cancer%20has%20four%20primary,(HR%2FHER2%2Dnegative)

# Methods

## Study Population

## Variables

## Statistical Analysis

# Results

## Clinicopathological characteristics

## Univariate Analysis

## Multivariate Analysis

# Discussion

### 2024-11-25 code/analysis/loglog-plots.R

@@ -51,7 +51,7 @@ plot_n_stage <- ggsurvplot(
   fun="cloglog",
   ylab = "log(-log(S(t))",
   xlab = "log(Time)",
-  legend.labs = c("N0/1", "N2/3"),
+  legend.labs = c("N0", "N1", "N2", "N3"),
   legend.title = "N Stage:")
 
 plot_m_stage <- ggsurvplot(

### 2024-11-25 code/analysis/table1.R

@@ -20,6 +20,7 @@ box::use(
 seer <- readRDS("data/processed/seer_tab.rds")
 attach(seer)
 
+# create table 1
 t1 <-
   seer |>
   gtsummary::tbl_summary(
@@ -27,7 +28,7 @@ t1 <-
     include = c(agecat, race, marriage_status, n_stage, m_stage,
                 gradecat, radiation, chemotherapy, surgery),
     label = list(
-      agecat ~ "Age (years)",
+      agecat ~ "Age at diagnosis",
       race ~ "Race",
       marriage_status ~ "Marital Status",
       n_stage ~ "N Stage",
@@ -48,7 +49,6 @@ t1 <-
     ),
     digits = c(agecat, race, marriage_status, n_stage, m_stage,
                gradecat, radiation, chemotherapy, surgery) ~ c(0, 1),
-    statistic_header = "**Statistic**"
   ) |>
   gtsummary::add_overall() |>
   gtsummary::add_p() |>
@@ -57,17 +57,10 @@ t1 <-
   ) |>
   gtsummary::as_gt()
 
+# save to png
 gt::gtsave(
   data = t1,
   filename = here::here("results/tables", "t1.png"),
   vwidth = 2400,
   vheight = 1350
 )
-
-t4b |>
-  gt::rm_caption() |>
-  gt::gtsave(
-    filename = here::here("results", "paper-t1.png"),
-    vwidth = 2400,
-    vheight = 1350
-  )

### 2024-11-25 code/analysis/table2.R

@@ -1,0 +1,89 @@
+# Table 2: Survival for IBC by Clinicopathological Characteristics
+
+library(survival)
+library(gtsummary)
+library(dplyr)
+
+# load data
+seer <- readRDS("data/processed/seer.rds")
+attach(seer)
+
+# 1, 5, and 10 year survival
+t2 <- tbl_survfit(
+  seer,
+  y = "Surv(survival_months, survival_status)",
+  include = c(subtype, age, race, marriage_status, n_stage, m_stage,
+              grade, radiation, chemotherapy, surgery),
+  label = list(
+    subtype ~ "Subtype",
+    age ~ "Age at diagnosis",
+    race ~ "Race",
+    marriage_status ~ "Marital Status",
+    n_stage ~ "N Stage",
+    m_stage ~ "M Stage",
+    grade ~ "Grade",
+    radiation ~ "Radiation",
+    chemotherapy ~ "Chemotherapy",
+    surgery ~ "Surgery"),
+  statistic = "{estimate}",
+  times = c(0, 12, 60, 120)
+  ) |>
+  gtsummary::modify_caption(
+    "Table 2: Survival for IBC by Clinicopathological Characteristics"
+  )
+
+# create and assign more descriptive labels
+labels = c(
+  "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
+  "Age at diagnosis", "20-49 years", "50-69 years", ">=70 years",
+  "Race", "White", "Black", "Other",
+  "Marital Status", "Unmarried", "Married",
+  "N Stage", "N0", "N1", "N2", "N3",
+  "M Stage", "M0", "M1",
+  "Grade", "I/II", "III/IV",
+  "Radiation", "No/Unknown", "Yes",
+  "Chemotherapy", "No/Unknown", "Yes",
+  "Surgery", "No/Unknown", "Yes")
+
+t2$table_body$label[1:36] <- labels
+t2$table_styling$header$label[5:8] <- c("**Median Survival (years)**", "**1 Year**", "**5 Year**", "**10 Year**")
+
+# since we can't do both median survival and specific times,
+# create a seperate table with median survival
+t2_med <- tbl_survfit(
+  seer,
+  y = "Surv(survival_months, survival_status)",
+  include = c(subtype, age, race, marriage_status, n_stage, m_stage,
+              grade, radiation, chemotherapy, surgery),
+  label = list(
[…]

### 2024-11-25 code/analysis/table3.R

@@ -1,0 +1,94 @@
+# Table 3: Cox Proportional Hazards Model Regression Results
+
+# load data
+seer <- readRDS("data/processed/seer.rds")
+attach(seer)
+
+# first univariate table
+t3_uv <- gtsummary::tbl_uvregression(
+  data = seer,
+  method = coxph,
+  y = Surv(survival_months, survival_status),
+  exponentiate = TRUE,
+  include = c(subtype, age, race, marriage_status, n_stage, m_stage, grade, radiation, chemotherapy, surgery),
+  label = list(
+    subtype ~ "Subtype", 
+    age ~ "Age",
+    race ~ "Race",
+    marriage_status ~ "Marital Status", 
+    n_stage ~ "N Stage", 
+    m_stage ~ "M Stage",
+    grade ~ "Grade",
+    radiation ~ "Radiation", 
+    chemotherapy ~ "Chemotherapy",
+    surgery ~ "Surgery"
+  ),
+  hide_n = TRUE,
+)
+
+# create and assign more descriptive labels
+labels = c(
+  "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
+  "Age at diagnosis", "20-49 years", "50-69 years", ">=70 years",
+  "Race", "White", "Black", "Other",
+  "Marital Status", "Unmarried", "Married",
+  "N Stage", "N0", "N1", "N2", "N3",
+  "M Stage", "M0", "M1",
+  "Grade", "I/II", "III/IV",
+  "Radiation", "No/Unknown", "Yes",
+  "Chemotherapy", "No/Unknown", "Yes",
+  "Surgery", "No/Unknown", "Yes")
+
+t3_uv$table_body$label[1:36] <- labels
+
+# next multivariable table
+t3_mv <- coxph(
+  Surv(survival_months, survival_status) ~ 
+    subtype + 
+    age + 
+    race + 
+    marriage_status + 
+    n_stage + 
+    m_stage + 
+    grade + 
+    radiation + 
+    chemotherapy + 
+    surgery,
+  data = seer, 
+  method = "breslow") |>
+  gtsummary::tbl_regression(
+    exp = TRUE,
+    label = list(
+      subtype ~ "Subtype", 
+      age ~ "Age",
+      race ~ "Race",
+      marriage_status ~ "Marital Status", 
+      n_stage ~ "N Stage", 
+      m_stage ~ "M Stage",
[…]

### 2024-11-25 code/processing/processing-script.R

@@ -191,7 +191,7 @@ seer_complete <- seer_clean %>%
   )
 
 # there are several variables that need to be categorized for analysis and/filtered
-# for this df leave the category names legible for creating tables
+# for this df leave the category names legible for creating table 1
 seer_catclean <- seer_complete %>%
   mutate(subtype = factor(
     subtype, levels = c("Luminal A", "Luminal B", "HER2 Positive", "Triple Negative")
@@ -219,14 +219,9 @@ seer_catclean <- seer_complete %>%
     grade %in% c(1, 2) ~ "I/II",
     grade %in% c(3, 4) ~ "III/IV",
     TRUE ~ NA_character_
-  )) %>%
-  mutate(n_stage = case_when(
-    n_stage %in% c("N0", "N1") ~ "N0/N1",
-    n_stage %in% c("N2", "N3") ~ "N2/N3",
-    TRUE ~ NA_character_
   ))
 
-# save this data with named categories as it will be useful for creating tables
+# save this data with named categories as it will be useful for creating table 1
 seer_tab <- seer_catclean %>%
   select(
     agecat,
@@ -248,123 +243,85 @@ saveRDS(seer_tab, file = "data/processed/seer_tab.rds")
 
 # there are several variables that need to be categorized for analysis and/filtered
 seer_cat <- seer_complete %>%
-  mutate(agecat = case_when(
-    age %in% c("20-24 years", "25-29 years", "30-34 years", "35-39 years", "40-44 years", "45-49 years") ~ 0,
-    age %in% c("50-54 years", "55-59 years", "60-64 years", "65-69 years") ~ 1,
-    age %in% c("70-74 years", "75-79 years", "80-84 years", "85+ years") ~ 2,
-    TRUE ~ NA_real_  # Keeps other values as NA if they don’t match any of the above groups
-  )) %>%
-  mutate(agecat0 = case_when(
-    agecat == 0 ~ 1,
-    TRUE ~ 0
-  )) %>%
-  mutate(agecat1 = case_when(
-    agecat == 1 ~ 1,
-    TRUE ~ 0
-  )) %>%
-  mutate(agecat2 = case_when(
-    agecat == 2 ~ 1,
[…]

### 2024-11-25 products/manuscript/manuscript.qmd
manuscript start

@@
 format: pdf
+bibliography:
+  - "refs.bib"
 ---
@@
 
-Inflammatory breast carcinoma (IBC) is a rare but aggressive subtype of breast cancer.
-Reportedly, IBC accounts for only 1% to 6% of breast cancer cases in the United States, yet it was responsible for 10% of all types of breast cancer-related deaths (Zhu)
+Inflammatory breast cancer (IBC) is a rare but aggressive subtype of breast cancer, accounting for 2% to 6% of all breast cancer cases in the United States [@devi2019perspectives; @hance2005trends].
+IBC presents differently than other types of breast cancer.
+It often lacks a breast lump and may not appear on a mammogram, making accurate diagnosis challenging and potentially delaying treatment.
+IBC causes symptoms of breast inflammation, such as swelling and redness, due to cancer cells blocking lymphatic vessels in the skin, giving the breast an "inflamed" appearance [@menta2018inflammatory].
 
-Inflammatory breast cancer (IBC) is a rare and aggressive variant of stage IIID breast cancer, with increased likelihood of metastasis upon diagnosis relative to non-inflammatory breast cancer.
-(Abraham)
+Clinically, IBC is categorized as T4d in the TNM classification for its aggressive biological behavior.
+In spite of multimodal therapy, IBC has a local recurrence rate of up to 50% and survival rates of only 35% to 40%, significantly lower than those of other types of breast cancer [@robertson2010inflammatory].
+Despite its rarity, IBC disproportionately accounts for up to 10% of all breast cancer-related deaths [@devi2019perspectives; @hance2005trends].
 
-Inflammatory breast cancer (IBC) is undoubtedly a unique biologic subtype of breast cancer with poor prognosis, which accounts for 1e5% of all breast cancer and comprises 7.0% of all breast cancerspecific mortality.
-(Pan)
+Because IBC is rare, the biological characteristics of IBC are not commonly reported and the molecular alterations that result in poor prognosis are not well understood.
+However, hormone receptors (HR) and human epidermal growth factor receptor-2 (HER2), which define the molecular status of IBC through immunohistochemistry (IHC), are fundamental markers used to demonstrate molecular features, predict prognosis, and optimize therapeutic regimens [@li2017outcomes].
+IBC tumors are characterized by combinations of hormone receptors and oncogenes, allowing the majority to be characterized as HR+/HER2- (Luminal A), HR+/HER+ (Luminal B), HR-/HER2+ (HER2 Positive), and Triple Negative by IHC.[@TCR28003].
+Studies have shown that these four molecular subtypes reveal the prognostic discrepancies in both common breast cancers and IBC [@robertson2010inflammatory; @le2015future; @tolaney2015adjuvant; @iwamoto2011different].
+A previous study examined the association between molecular subtypes and survival outcomes in IBC patients from 2010 to 2013 [@li2017outcomes].
+However, no updated population-based studies have investigated this relationship.
 
-According to the TNM classification, IBC is classified as T4d (a category reserved for IBC) for its aggressive biological behaviour.
-With great efforts to improve survival, including advancements in surgical technique, targeted therapy, radiotherapy and chemotherapy and improvements in supportive treatment, the prognosis of IBC has increased (Pan)
-
-Although IBC demonstrates an inflammatory appearance, it is caused by cancer cells obstructing the dermal lymphatics of the breasts. (Zhu)
-
-IBC is categorized as T4d (a category designated for IBC) in the TNM classification with its biologically aggressive character.
-With the application of a multimodal therapy strategy that includes systemic neoadjuvant chemotherapy, surgery, and radiotherapy, the clinical outcomes of IBC have significantly improved.
-(Zhu)
-
-However, the approximately 35% to 40% 5-year cumulative survival rate is much lower compared with other breast cancers. (Zhu)
-
-Historically, single-modality treatment to cure IBC was not successful; 90% of patients developed recurrent and/or metastatic disease within 2 years, and the 5-year survival rate was 5%.
-Combinations of neoadjuvant systemic chemotherapy, surgery, and radiation therapy have led to an improved prognosis.
-However, the overall 5-year survival rate for patients with IBC is still very low, at 30% (Yamauchi)
-
-The aim of this paper is to determine the effect of cancer subtype on survival in adult femaes with IBC.
-
-"T4 (includes T4a, T4b, T4c, and T4d): Tumor of any size growing into the chest wall or skin.
-This includes inflammatory breast cancer." https://www.cancer.org/cancer/types/breast-cancer/understanding-a-breast-cancer-diagnosis/stages-of-breast-cancer.html#:~:text=T1%20(includes%20T1a%2C%20T1b%2C,the%20chest%20wall%20or%20skin.
-
-"Inflammatory breast cancer (IBC) is rare.
-It accounts for only 1% to 5% of all breast cancers.
-Although it is a type of invasive ductal carcinoma, its symptoms, outlook, and treatment are different.
-IBC causes symptoms of breast inflammation like swelling and redness, which is caused by cancer cells blocking lymph vessels in the skin causing the breast to look "inflamed." " https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html
-
-"IBC doesn't look like a typical breast cancer.
-It often does not cause a breast lump, and it might not show up on a mammogram.
-This makes it harder to diagnose., IBC tends to occur in younger women (younger than 40 years of age)., Black women appear to develop IBC more often than White women., IBC is more common among women who are overweight or obese.
-IBC tends to be more aggressive—it grows and spreads much more quickly—than more common types of breast cancer., IBC is always at least at a locally advanced stage when it’s first diagnosed because the breast cancer cells have grown into the skin.
-(This means it is at least stage III.), In about 1 of 3 cases, IBC has already spread (metastasized) to distant parts of the body when it is diagnosed.
-This makes it harder to treat successfully., Women with IBC tend to have a worse prognosis (outcome) than women with other common types of breast cancer." https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html
-
-"All inflammatory breast cancers start as stage III (T4dNXM0) since they involve the skin.
-If the cancer has spread outside the breast to distant parts of the body, it is stage IV." https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html
-
-"Inflammatory breast cancer (IBC) that has not spread outside the breast is stage III.
-In most cases, treatment is chemotherapy first to try to shrink the tumor, followed by surgery to remove the cancer.
-Radiation and often other treatments, like more chemotherapy or targeted drug therapy, are given after surgery.
-Because IBC is so aggressive, breast conserving surgery (lumpectomy) and sentinel lymph node biopsy are typically not part of the treatment.
-IBC that has spread to other parts of the body (stage IV) may be treated with chemotherapy, hormone therapy, and/or targeted drugs." https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/inflammatory-breast-cancer.html
-
-"The most common and widely accepted classification of breast cancer is from an immunohistochemical perspective, based on the expression of the following hormone receptors: estrogen (ER), progesterone (PR) and human epidermal growth factor (HER2).
-Accordingly, the following four subtypes of breast cancer are widely recognized: luminal A, luminal B, HER2-positive, and triple-negative.
-" https://www.ncbi.nlm.nih.gov/books/NBK583808/#:~:text=Accordingly%2C%20the%20following%20four%20subtypes,positive%2C%20and%20triple%2Dnegative.
-
-"Breast cancer has four primary molecular subtypes, defined in large part by hormone receptors (HR) and other types of proteins involved (or not involved) in each cancer: Luminal A or HR+/HER2- (HR-positive/HER2-negative) Luminal B or HR+/HER2+ (HR-positive/HER2-positive) Triple-negative or HR-/HER2- (HR/HER2-negative) HER2-positive" https://www.cancercenter.com/cancer-types/breast-cancer/types/breast-cancer-molecular-types#:~:text=Breast%20cancer%20has%20four%20primary,(HR%2FHER2%2Dnegative)
+Therefore, using the most recent Surveillance, Epidemiology, and End Results (SEER) data, this study aims to assess the effect of IBC molecular subtype on breast cancer-specific mortality in adult females with IBC.
 
@@
 
+The National Cancer Institute's Surveillance, Epidemiology, and End Results (SEER) database is an authoritative source of information on cancer incidence and survival in the United States.
+SEER currently collects and publishes cancer incidence and survival data from population-based cancer registries covering approximately 48% of the U.S.
+population.
+This study utilized SEER incidence data from 2010 to 2021.
+
 ## Study Population
 
+The study population focuses on adult females diagnosed with IBC.
+Data was extracted from the SEER 17 Registries (2000-2021) using the SEER\*Stat software (version 8.4.4; Surveillance Research Program, NCI, Bethesda, MD).
+The specific criteria used to identify patients with IBC were as follows: (1) Site and Morphology (TNM 7/CS v0204 Schema recode) was limited to "Breast"; (2) Derived TNM classification, AJCC 7th edition (2010-2015) was limited to "T4d", SEER Combined T (2016-2017) was limited to "c2D" and "p4D", and EOD 8th edition (2018+) was limited to "T4D"; and (3) Histology type (ICD-O-3) ranged from 8500 to 8549.
+Year of diagnosis was limited to 2010 to 2021 because SEER started collecting breast cancer molecular subtype in 2010.
+To limit our data to adult females, age at diagnosis was limited to patents 20 years and older and sex was limited to "Female".
+
 ## Variables
+
+For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer dx), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer dx), Dead (missing/unknown COD)).
+In addition, cases were excluded from the study if they had incomplete information on any of these characteristics.
+Patients who survived less than one month were also excluded.
+To limit our analysis to breast cancer-specific mortality, patients with unknown cause of death or patients with cause of death to another casue were excluded.
 

### 2024-11-25 products/manuscript/manuscript.qmd
manuscript updates

@@
 
-**Keywords:**
-
 \pagebreak
@@
 However, hormone receptors (HR) and human epidermal growth factor receptor-2 (HER2), which define the molecular status of IBC through immunohistochemistry (IHC), are fundamental markers used to demonstrate molecular features, predict prognosis, and optimize therapeutic regimens [@li2017outcomes].
-IBC tumors are characterized by combinations of hormone receptors and oncogenes, allowing the majority to be characterized as HR+/HER2- (Luminal A), HR+/HER+ (Luminal B), HR-/HER2+ (HER2 Positive), and Triple Negative by IHC.[@TCR28003].
+IBC tumors are characterized by combinations of hormone receptors and oncogenes, allowing the majority to be characterized as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 Positive (HR-/HER2+), and Triple Negative by IHC.[@TCR28003].
 Studies have shown that these four molecular subtypes reveal the prognostic discrepancies in both common breast cancers and IBC [@robertson2010inflammatory; @le2015future; @tolaney2015adjuvant; @iwamoto2011different].
@@
 
-The National Cancer Institute's Surveillance, Epidemiology, and End Results (SEER) database is an authoritative source of information on cancer incidence and survival in the United States.
+The National Cancer Institute's Surveillance, Epidemiology, and End Results (SEER) database is an authoritative source of information on cancer incidence and survival in the United States (https://seer.cancer.gov/).
 SEER currently collects and publishes cancer incidence and survival data from population-based cancer registries covering approximately 48% of the U.S.
@@
 
-For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer dx), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer dx), Dead (missing/unknown COD)).
-In addition, cases were excluded from the study if they had incomplete information on any of these characteristics.
-Patients who survived less than one month were also excluded.
-To limit our analysis to breast cancer-specific mortality, patients with unknown cause of death or patients with cause of death to another casue were excluded.
+For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer), Dead (missing/unknown COD)).
+Breast cancer subtypes were classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive (HR-/HER2+), and Triple Negative.
+In addition, cases with incomplete information on any of these characteristics and patients who survived less than one month were excluded from the study.
+Additionally, to focus on breast cancer-specific mortality, patients with an unknown cause of death or whose death was attributed to other causes were excluded.
 
 ## Statistical Analysis
+
+Clinicopatholical characteristics were compared between IBC subtype using Pearson's Chi-squared test.
+Median, 1-year, 5-year, and 10-year survival for IBC is estimated for each variable in the analysis using the Kaplain-Meier method and variables were compared using the log-rank test in the subgroups.
+Univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
+We obtained hazard ratios (HR) and 95% confidence intervals for each covariate.
+All statistical analyses were performed using the survival package in R version 4.3.3.
 
@@
 ## Clinicopathological characteristics
+
+A total of 5969 adult female IBC cases were eligible during 2010-2021 in the SEER database for our study population.
+Of these, 1829 were excluded; 165 had a survival time less than 1 month, 487 had an unknown or competing cause of death, and 1177 had incomplete information.
+The remaining 4140 were included in our analysis, as shown in Table 1.
+The clinicopathological characteristics of the eligible patients are shown in Table 2.
+
+## Survival Analysis
 

### 2024-11-25 products/manuscript/manuscript.tex
manuscript updates

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={epid 8010 manuscript},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{epid 8010 manuscript}
\author{}
\date{}

\begin{document}
\maketitle

\textbf{Authors}

\begin{itemize}
\tightlist
\item
  Murphy John\(^{1}\)
\end{itemize}

\textbf{Author affiliation}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Department of Epidemiology and Biostatistics, University of Georgia,
  Athens, GA, USA.
\end{enumerate}

\pagebreak

\section{Abstract}\label{abstract}

\textbf{Background:}

\textbf{Methods:}

\textbf{Results:}

\textbf{Conclusions:}

\pagebreak

\section{Introduction}\label{introduction}

Inflammatory breast cancer (IBC) is a rare but aggressive subtype of
breast cancer, accounting for 2\% to 6\% of all breast cancer cases in
the United States (Devi et al. 2019; Hance et al. 2005). IBC presents
differently than other types of breast cancer. It often lacks a breast
lump and may not appear on a mammogram, making accurate diagnosis
challenging and potentially delaying treatment. IBC causes symptoms of
breast inflammation, such as swelling and redness, due to cancer cells
blocking lymphatic vessels in the skin, giving the breast an
``inflamed'' appearance (Menta et al. 2018).

Clinically, IBC is categorized as T4d in the TNM classification for its
aggressive biological behavior. In spite of multimodal therapy, IBC has
a local recurrence rate of up to 50\% and survival rates of only 35\% to
40\%, significantly lower than those of other types of breast cancer
(Robertson et al. 2010). Despite its rarity, IBC disproportionately
accounts for up to 10\% of all breast cancer-related deaths (Devi et al.
2019; Hance et al. 2005).

Because IBC is rare, the biological characteristics of IBC are not
commonly reported and the molecular alterations that result in poor
prognosis are not well understood. However, hormone receptors (HR) and
human epidermal growth factor receptor-2 (HER2), which define the
molecular status of IBC through immunohistochemistry (IHC), are
fundamental markers used to demonstrate molecular features, predict
prognosis, and optimize therapeutic regimens (Li et al. 2017). IBC
tumors are characterized by combinations of hormone receptors and
oncogenes, allowing the majority to be characterized as Luminal A
(HR+/HER2-), Luminal B (HR+/HER+), HER2 Positive (HR-/HER2+), and Triple
Negative by IHC.(Bonito, Cantile, and Botti 2019). Studies have shown
that these four molecular subtypes reveal the prognostic discrepancies
in both common breast cancers and IBC (Robertson et al. 2010; Le Du et
al. 2015; Tolaney et al. 2015; Iwamoto et al. 2011). A previous study
examined the association between molecular subtypes and survival
outcomes in IBC patients from 2010 to 2013 (Li et al. 2017). However, no
updated population-based studies have investigated this relationship.

Therefore, using the most recent Surveillance, Epidemiology, and End
Results (SEER) data, this study aims to assess the effect of IBC
molecular subtype on breast cancer-specific mortality in adult females
with IBC.

\section{Methods}\label{methods}

The National Cancer Institute's Surveillance, Epidemiology, and End
Results (SEER) database is an authoritative source of information on
cancer incidence and survival in the United States
(https://seer.cancer.gov/). SEER currently collects and publishes cancer
incidence and survival data from population-based cancer registries
covering approximately 48\% of the U.S. population. This study utilized
SEER incidence data from 2010 to 2021.

\subsection{Study Population}\label{study-population}

The study population focuses on adult females diagnosed with IBC. Data
was extracted from the SEER 17 Registries (2000-2021) using the
SEER*Stat software (version 8.4.4; Surveillance Research Program, NCI,
Bethesda, MD). The specific criteria used to identify patients with IBC
were as follows: (1) Site and Morphology (TNM 7/CS v0204 Schema recode)
was limited to ``Breast''; (2) Derived TNM classification, AJCC 7th
edition (2010-2015) was limited to ``T4d'', SEER Combined T (2016-2017)
was limited to ``c2D'' and ``p4D'', and EOD 8th edition (2018+) was
limited to ``T4D''; and (3) Histology type (ICD-O-3) ranged from 8500 to
8549. Year of diagnosis was limited to 2010 to 2021 because SEER started
collecting breast cancer molecular subtype in 2010. To limit our data to
adult females, age at diagnosis was limited to patents 20 years and
older and sex was limited to ``Female''.

\subsection{Variables}\label{variables}

For each case, the following information from SEER was obtained: age at
diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other),
marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage
(M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes),
chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival
months, cause-specific death classification (Alive or dead of other
cause, Dead (attributable to this cancer), Dead (missing/unknown COD)),
and other cause of death classification (Alive or dead due to cancer,
Dead (attributable to causes other than this cancer), Dead
(missing/unknown COD)). Breast cancer subtypes were classified as
Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive (HR-/HER2+),
and Triple Negative. In addition, cases with incomplete information on
any of these characteristics and patients who survived less than one
month were excluded from the study. Additionally, to focus on breast
cancer-specific mortality, patients with an unknown cause of death or
whose death was attributed to other causes were excluded.

\subsection{Statistical Analysis}\label{statistical-analysis}

Clinicopatholical characteristics were compared between IBC subtype
using Pearson's Chi-squared test. Median, 1-year, 5-year, and 10-year
survival for IBC is estimated for each variable in the analysis using
the Kaplain-Meier method and variables were compared using the log-rank
test in the subgroups. Univariate and multivariate Cox proportional
hazards regressions were used to model the relationship between
potential covariates and IBC-specific survival. We obtained hazard
ratios (HR) and 95\% confidence intervals for each covariate. All
statistical analyses were performed using the survival package in R
version 4.3.3.

\section{Results}\label{results}

\subsection{Clinicopathological
characteristics}\label{clinicopathological-characteristics}

A total of 5969 adult female IBC cases were eligible during 2010-2021 in
the SEER database for our study population. Of these, 1829 were
excluded; 165 had a survival time less than 1 month, 487 had an unknown
or competing cause of death, and 1177 had incomplete information. The
remaining 4140 were included in our analysis, as shown in Table 1. The
clinicopathological characteristics of the eligible patients are shown
in Table 2.

\subsection{Survival Analysis}\label{survival-analysis}

\subsection{Univariate Analysis}\label{univariate-analysis}

\subsection{Multivariate Analysis}\label{multivariate-analysis}

\section*{Discussion}\label{discussion}
\addcontentsline{toc}{section}{Discussion}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-TCR28003}
Bonito, Maurizio Di, Monica Cantile, and Gerardo Botti. 2019.
{``Pathological and Molecular Characteristics of Inflammatory Breast
Cancer.''} \emph{Translational Cancer Research} 8 (Suppl 5).
\url{https://tcr.amegroups.org/article/view/28003}.

\bibitem[\citeproctext]{ref-devi2019perspectives}
Devi, Gayathri R, Holly Hough, Nadine Barrett, Massimo Cristofanilli,
Beth Overmoyer, Neil Spector, Naoto T Ueno, et al. 2019. {``Perspectives
on Inflammatory Breast Cancer (IBC) Research, Clinical Management and
Community Engagement from the Duke IBC Consortium.''} \emph{Journal of
Cancer} 10 (15): 3344.

\bibitem[\citeproctext]{ref-hance2005trends}
Hance, Kenneth W, William F Anderson, Susan S Devesa, Heather A Young,
and Paul H Levine. 2005. {``Trends in Inflammatory Breast Carcinoma
Incidence and Survival: The Surveillance, Epidemiology, and End Results
Program at the National Cancer Institute.''} \emph{Journal of the
National Cancer Institute} 97 (13): 966--75.

\bibitem[\citeproctext]{ref-iwamoto2011different}
Iwamoto, Takayuki, Giampaolo Bianchini, Yuan Qi, Massimo Cristofanilli,
Anthony Lucci, Wendy A Woodward, James M Reuben, et al. 2011.
{``Different Gene Expressions Are Associated with the Different
Molecular Subtypes of Inflammatory Breast Cancer.''} \emph{Breast Cancer
Research and Treatment} 125: 785--95.

\bibitem[\citeproctext]{ref-le2015future}
Le Du, Fanny, Bedrich L Eckhardt, Bora Lim, Jennifer K Litton, Stacy
Moulder, Funda Meric-Bernstam, Ana M Gonzalez-Angulo, and Naoto T Ueno.
2015. {``Is the Future of Personalized Therapy in Triple-Negative Breast
Cancer Based on Molecular Subtype?''} \emph{Oncotarget} 6 (15): 12890.

\bibitem[\citeproctext]{ref-li2017outcomes}
Li, Juanjuan, Yue Xia, Qi Wu, Shan Zhu, Chuang Chen, Wen Yang, Wen Wei,
and Shengrong Sun. 2017. {``Outcomes of Patients with Inflammatory
Breast Cancer by Hormone Receptor-and HER2-Defined Molecular Subtypes: A
Population-Based Study from the SEER Program.''} \emph{Oncotarget} 8
(30): 49370.

\bibitem[\citeproctext]{ref-menta2018inflammatory}
Menta, Arjun, Tamer M Fouad, Anthony Lucci, Huong Le-Petross, Michael C
Stauder, Wendy A Woodward, Naoto T Ueno, and Bora Lim. 2018.
{``Inflammatory Breast Cancer: What to Know about This Unique,
Aggressive Breast Cancer.''} \emph{Surgical Clinics} 98 (4): 787--800.

\bibitem[\citeproctext]{ref-robertson2010inflammatory}
Robertson, Fredika M, Melissa Bondy, Wei Yang, Hideko Yamauchi, Shannon
Wiggins, Samira Kamrudin, Savitri Krishnamurthy, et al. 2010.
{``Inflammatory Breast Cancer: The Disease, the Biology, the
Treatment.''} \emph{CA: A Cancer Journal for Clinicians} 60 (6):
351--75.

\bibitem[\citeproctext]{ref-tolaney2015adjuvant}
Tolaney, Sara M, William T Barry, Chau T Dang, Denise A Yardley, Beverly
Moy, P Kelly Marcom, Kathy S Albain, et al. 2015. {``Adjuvant Paclitaxel
and Trastuzumab for Node-Negative, HER2-Positive Breast Cancer.''}
\emph{New England Journal of Medicine} 372 (2): 134--41.

\end{CSLReferences}

\end{document}

### 2024-11-25 code/analysis/flowchart.R

@@ -7,25 +7,21 @@ library(grid)
 library(magrittr)
 
 
-eligible <- boxGrob(glue("Breast cencer patients diagnosed with \ninflammatory breast carcinoma between 2010-2021",
+eligible <- boxGrob(glue("Adult female patients diagnosed with \ninflammatory breast carcinoma between 2010-2021",
                          "N = {pop}",
-                         pop = txtInt(5986),
+                         pop = txtInt(5969),
                          .sep = "\n"))
 
-included <- boxGrob(glue("Eligible inflammatory breast \ncancer patients",
+included <- boxGrob(glue("Eligible adult female inflammatory \nbreast cancer patients",
                          "n = {incl}",
                          incl = txtInt(4140),
                          .sep = "\n"))
 
 excluded <- boxGrob(glue("Excluded (n = {tot}):",
-                         " - Age at diagnosis < 20 years: {age}",
-                         " - Male sex: {male}",
                          " - Survival time < 1 month: {survival}",
                          " - Unknown or competing \ncause of death: {cod}",
                          " - Incomplete or unknown \nclinicopathological characteristics: {incomplete}",
-                         tot = txtInt(1846),
-                         age = 1,
-                         male = 16,
+                         tot = txtInt(1829),
                          survival = 165,
                          cod = 487,
                          incomplete = txtInt(1177),

### 2024-11-25 code/analysis/table2.R

@@ -28,8 +28,9 @@ t2 <- tbl_survfit(
   statistic = "{estimate}",
   times = c(0, 12, 60, 120)
   ) |>
+  add_p() |>
   gtsummary::modify_caption(
-    "Table 2: Survival for IBC by Clinicopathological Characteristics"
+    "Table 2: Survival for IBC by Demographics <br> and Clinicopathologic Characteristics"
   )
 
 # create and assign more descriptive labels
@@ -46,7 +47,7 @@ labels = c(
   "Surgery", "No/Unknown", "Yes")
 
 t2$table_body$label[1:36] <- labels
-t2$table_styling$header$label[5:8] <- c("**Median Survival (years)**", "**1 Year**", "**5 Year**", "**10 Year**")
+t2$table_styling$header$label[6:9] <- c("**Median Survival (years)**", "**1 Year**", "**5 Year**", "**10 Year**")
 
 # since we can't do both median survival and specific times,
 # create a seperate table with median survival

### 2024-11-26 products/manuscript/manuscript.qmd
writing

@@
 
-**Authors**
+**Author**
 
--   Murphy John$^{1}$
+Murphy John$^{1}$
 
@@
 
-For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer), Dead (missing/unknown COD)).
+For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer), Dead (missing/unknown COD)).
 Breast cancer subtypes were classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive (HR-/HER2+), and Triple Negative.
@@
 Median, 1-year, 5-year, and 10-year survival for IBC is estimated for each variable in the analysis using the Kaplain-Meier method and variables were compared using the log-rank test in the subgroups.
+
+### Proportional Hazards Assumption
+
 Univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
@@
 Of these, 1829 were excluded; 165 had a survival time less than 1 month, 487 had an unknown or competing cause of death, and 1177 had incomplete information.
-The remaining 4140 were included in our analysis, as shown in Table 1.
-The clinicopathological characteristics of the eligible patients are shown in Table 2.
+The remaining 4140 were included in our analysis, as shown in Figure 1.
+
+```{r}
+#| label: fig-flowchart
+#| fig-cap: |
+#|   [Flowchart of IBC patients screening on SEER database]{color="#FF2400"}
+#| out.width: 100%
+#| out.height: 80%
+knitr::include_graphics(
+	here::here("results", "figures", "flowchart.png")
+)
+```
+
+The clinicopathological characteristics of eligible patients are summarized in Table 1.
+The age subgroup of 20–49 years accounts for 35.1% of Luminal B cases, a higher proportion compared to other subtypes.
+Moreover, Luminal B and HER2-positive subtypes appear to affect younger age groups more frequently than Luminal A, with 85.4% and 84.5% of cases, respectively, occurring in patients under 69 years, compared to 76.5% for Luminal A.
+A higher proportion of Black patients is observed in the Triple-Negative group compared to other subtypes.
+Marital status analysis shows that the Triple-Negative group has the highest percentage of unmarried women.
+Triple-Negative cases also have the highest percentage of N3 node stage and exhibit the greatest proportion of grade III/IV cases.
+
+```{r}
+#| label: tbl-table1
+#| tbl-cap: |
+#|   [Clinicopathological characteristics by cancer subtype]{color="#FF2400"}
+knitr::include_graphics(
+	here::here("results", "tables", "t1.png")
+)
+```
 
@@
 
+Table 2 presents the median survival in years, along with 1-, 5-, and 10-year survival percentages, stratified by clinicopathological characteristics, all of which show statistically significant differences based on the log-rank test.
+Survival outcomes differ significantly by subtype, with the Triple-Negative subtype showing the poorest prognosis, including a median survival of 1.7 years and 1-, 5-, and 10-year survival rates of 67%, 24%, and 21%, respectively.
+Conversely, the Luminal B subtype has the most favorable outcomes, with a median survival of 9.4 years and a 10-year survival rate of 50%.
+Racial disparities are evident, as Black patients have the poorest survival outcomes, with a median survival of 2.4 years compared to 4.6 years for White patients and 5.4 years for patients of other races.
+Unmarried women experience poorer survival compared to married women.
+Although 1-year survival is similar across all node stages, nodal involvement significantly affects long-term outcomes, with the N3 stage showing the lowest median survival of 2.7 years and the poorest 5- and 10-year survival rates.
+Survival differences are also pronounced by distant metastasis (M stage), with M1 stage associated with worse outcomes.
+Additionally, higher cancer grades correlate with poorer survival.
+In terms of treatment, patients who receive radiation, chemotherapy, or surgery exhibit better survival outcomes across all measurements.
+
+```{r}
+#| label: tbl-table2
+#| tbl-cap: |
+#|   [Survival for IBC by Clinicopathological Characteristics]{color="#FF2400"}
+knitr::include_graphics(
+	here::here("results", "tables", "t2.png")
+)
+```
+
+We conducted univariate and multivariate analyses based on the data in Table 2, with the results of the Cox proportional hazards regression presented in Table 3.
+
+```{r}
+#| label: tbl-table3
+#| tbl-cap: |
+#|   [Cox Proportional Hazards Regression
+#|   Model for Survival in Adult Women with IBC]{color="#FF2400"}
+knitr::include_graphics(
+	here::here("results", "tables", "t3.png")
+)
+```
+
 ## Univariate Analysis
+
+In the univariate analysis, subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were all significantly associated with IBC-specific survival.
+These variables were therefore included in the multivariate analysis.
 
@@
 
+The multivariate analysis revealed that subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were independent predictors of IBC-specific survival.
+The Luminal B subtype was associated with better survival (aHR=0.72, 95%CI: 0.56-0.74, P\<0.0001), whereas the Triple-Negative subtype showed the poorest survival outcomes (aHR=2.51, 95%CI: 1.79-2.21, P\<0.0001).
+In terms of treatment, our results find that the existing therapies are effective in improving survival, that is radiation (aHR=0.76, 95%CI: 0.69-0.85, P\<0.0001), chemotherapy (aHR=0.50, 95%CI: 0.44-0.57, P\<0.0001), and surgery (aHR=0.59, 95%CI: 0.52-0.66, P\<0.0001).
+
 # Discussion
+
+In this large population-based cohort of women diagnosed with IBC, we found a better survival in Luminal B molecular subtype and poorer in Triple Negative.
+We found that Found that Black patients made up the smallest percentage of Luminal B subtype and the greatest of Triple Negative.
+Similar findings were observed for patients who were unmarried at diagnosis.
+
+It would be interesting to look at the combination of treatments (radiation, chemo, surgery)

### 2024-11-26 code/analysis/table1.R

@@ -52,9 +52,6 @@ t1 <-
   ) |>
   gtsummary::add_overall() |>
   gtsummary::add_p() |>
-  gtsummary::modify_caption(
-    "Table 1: Clinicopathological characteristics by cancer subtype"
-  ) |>
   gtsummary::as_gt()
 
 # save to png

### 2024-11-26 code/analysis/table2.R

@@ -28,10 +28,7 @@ t2 <- tbl_survfit(
   statistic = "{estimate}",
   times = c(0, 12, 60, 120)
   ) |>
-  add_p() |>
-  gtsummary::modify_caption(
-    "Table 2: Survival for IBC by Demographics <br> and Clinicopathologic Characteristics"
-  )
+  add_p()
 
 # create and assign more descriptive labels
 labels = c(

### 2024-11-26 code/analysis/table3.R

@@ -80,11 +80,7 @@ t3 <-
   gtsummary::tbl_merge(
     list(t3_uv, t3_mv),
     tab_spanner = c("Crude", "Adjusted")
-  ) |>
-  gtsummary::modify_caption(
-    "Table 3: Cox Proportional Hazards Regression <br> Model for Survival in Adult Women with IBC"
   )
-t3
 
 gt::gtsave(
   data = t3 |> gtsummary::as_gt(),

### 2024-12-02 products/manuscript/manuscript.qmd
manuscript edits

@@
 
-For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer), Dead (missing/unknown COD)).
+For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N (node) stage (N0, N1, N2, N3), M (metastasis) stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer), Dead (missing/unknown COD)).
 Breast cancer subtypes were classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive (HR-/HER2+), and Triple Negative.
 In addition, cases with incomplete information on any of these characteristics and patients who survived less than one month were excluded from the study.
-Additionally, to focus on breast cancer-specific mortality, patients with an unknown cause of death or whose death was attributed to other causes were excluded.
+To focus on IBC-specific mortality, patients with an unknown cause of death or whose death was attributed to other causes were excluded.
 
@@
 
-Clinicopatholical characteristics were compared between IBC subtype using Pearson's Chi-squared test.
-Median, 1-year, 5-year, and 10-year survival for IBC is estimated for each variable in the analysis using the Kaplain-Meier method and variables were compared using the log-rank test in the subgroups.
-
-### Proportional Hazards Assumption
-
-Univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
+Clinicopathological characteristics were compared between IBC subtype using Pearson's Chi-squared test.
+Median, 1-year, 5-year, and 10-year survival for IBC is estimated for each variable in using the Kaplain-Meier method and compared using the log-rank test in the subgroups.
+The proportional hazards assumption was assessed visually with log-log plots and statistically by calculating Schoenfeld residuals and the Pearson's product moment correlation coefficient of the Schoenfeld residuals and ranked survival times.
+After determining the assumption of proportional hazard was met, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
 We obtained hazard ratios (HR) and 95% confidence intervals for each covariate.
@@
 ```
+
+## Proportional Hazards Assumption
+
+The proportional hazards assumption was assessed for each covariate visually using log-log plot, as shown in Figure 2.
+
+```{r}
+#| label: fig-flowchart
+#| fig-cap: |
+#|   [Log-log plots]{color="#FF2400"}
+#| out.width: 100%
+#| out.height: 80%
+knitr::include_graphics(
+	here::here("results", "figures", "loglog_plot.png")
+)
+```
+
+Upon inspection, there may be slight deviations from parallelity in the covariates subtype, radiation, chemotherapy, and surgery.
+Thus, the full Cox proportional hazards model is fit to the data and Schoenfeld residuals are extracted.
+The correlation between Schoenfeld residuals and ranked survival times for each covariate is shown in Table 2.
 

### 2024-12-02 code/final-project.Rmd
correlation table

@@
 
-# table 3: Correlation between Schoenfeld residuals and ranked survival time
-t3_raw <- cor_df %>%
-  add_row(Correlation = "", P_Value = "", .before = 1) %>%
-  add_row(Correlation = "", P_Value = "", .before = 1) %>%
-  add_row(Correlation = "", P_Value = "", .before = 6) %>%
-  add_row(Correlation = "", P_Value = "", .before = 6) %>%
-  add_row(Correlation = "", P_Value = "", .before = 10) %>%
-  add_row(Correlation = "", P_Value = "", .before = 10) %>%
-  add_row(Correlation = "", P_Value = "", .before = 14) %>%
-  add_row(Correlation = "", P_Value = "", .before = 14) %>%
-  add_row(Correlation = "", P_Value = "", .before = 17) %>%
-  add_row(Correlation = "", P_Value = "", .before = 17) %>%
-  add_row(Correlation = "", P_Value = "", .before = 20) %>%
-  add_row(Correlation = "", P_Value = "", .before = 20) %>%
-  add_row(Correlation = "", P_Value = "", .before = 23) %>%
-  add_row(Correlation = "", P_Value = "", .before = 23) %>%
-  add_row(Correlation = "", P_Value = "", .before = 26) %>%
-  add_row(Correlation = "", P_Value = "", .before = 26) %>%
-  add_row(Correlation = "", P_Value = "", .before = 29) %>%
-  add_row(Correlation = "", P_Value = "", .before = 29) %>%
-  add_row(Correlation = "", P_Value = "", .before = 32) %>%
-  add_row(Correlation = "", P_Value = "", .before = 32)
-
 ```

### 2024-12-02 products/manuscript/manuscript.qmd
correlation table

@@
 
-The proportional hazards assumption was assessed for each covariate visually using log-log plot, as shown in Figure 2.
+The proportional hazards assumption was assessed for each covariate visually using log minus log plots, as shown in Figure 2.
 
@@
 
-Upon inspection, there may be slight deviations from parallelity in the covariates subtype, radiation, chemotherapy, and surgery.
-Thus, the full Cox proportional hazards model is fit to the data and Schoenfeld residuals are extracted.
-The correlation between Schoenfeld residuals and ranked survival times for each covariate is shown in Table 2.
+Upon inspection, slight deviations from parallelism were observed for the covariates subtype, radiation, chemotherapy, and surgery.
+Therefore, the full Cox proportional hazards model was fitted to the data, and Schoenfeld residuals were extracted.
+The correlation between Schoenfeld residuals and ranked survival times for each covariate is presented in Table 2.
 

### 2024-12-02 code/analysis/table2.R

@@ -1,87 +1,47 @@
-# Table 2: Survival for IBC by Clinicopathological Characteristics
+# Table 2: Schoenfeld residual and ranked survival time correlations 
 
 library(survival)
-library(gtsummary)
 library(dplyr)
+library(gtsummary)
 
 # load data
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-# 1, 5, and 10 year survival
-t2 <- tbl_survfit(
-  seer,
-  y = "Surv(survival_months, survival_status)",
-  include = c(subtype, age, race, marriage_status, n_stage, m_stage,
-              grade, radiation, chemotherapy, surgery),
-  label = list(
-    subtype ~ "Subtype",
-    age ~ "Age at diagnosis",
-    race ~ "Race",
-    marriage_status ~ "Marital Status",
-    n_stage ~ "N Stage",
-    m_stage ~ "M Stage",
-    grade ~ "Grade",
-    radiation ~ "Radiation",
-    chemotherapy ~ "Chemotherapy",
-    surgery ~ "Surgery"),
-  statistic = "{estimate}",
-  times = c(0, 12, 60, 120)
-  ) |>
-  add_p()
+# fit the full cox proportional hazards model
+model <- coxph(
+  Surv(survival_months, survival_status) ~ 
+    relevel(factor(subtype), ref = '0') +
+    relevel(factor(age), ref = '0') + 
+    relevel(factor(race), ref = '0') +
+    relevel(factor(marriage_status),ref='9') + 
+    relevel(factor(n_stage),ref='0') + 
+    relevel(factor(m_stage),ref='0') +
+    relevel(factor(grade), ref='0') +
+    relevel(factor(radiation), ref='0') + 
+    relevel(factor(chemotherapy), ref='0') +
+    relevel(factor(surgery), ref='0'),
+  data = seer, 
+  method = "breslow")
 
-# create and assign more descriptive labels
-labels = c(
-  "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
-  "Age at diagnosis", "20-49 years", "50-69 years", ">=70 years",
-  "Race", "White", "Black", "Other",
-  "Marital Status", "Unmarried", "Married",
-  "N Stage", "N0", "N1", "N2", "N3",
-  "M Stage", "M0", "M1",
-  "Grade", "I/II", "III/IV",
[…]

### 2024-12-02 code/analysis/table3.R

@@ -1,30 +1,34 @@
-# Table 3: Cox Proportional Hazards Model Regression Results
+# Table 3: Survival for IBC by Clinicopathological Characteristics
+
+library(survival)
+library(gtsummary)
+library(dplyr)
 
 # load data
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-# first univariate table
-t3_uv <- gtsummary::tbl_uvregression(
-  data = seer,
-  method = coxph,
-  y = Surv(survival_months, survival_status),
-  exponentiate = TRUE,
-  include = c(subtype, age, race, marriage_status, n_stage, m_stage, grade, radiation, chemotherapy, surgery),
+# 1, 5, and 10 year survival
+t3 <- tbl_survfit(
+  seer,
+  y = "Surv(survival_months, survival_status)",
+  include = c(subtype, age, race, marriage_status, n_stage, m_stage,
+              grade, radiation, chemotherapy, surgery),
   label = list(
-    subtype ~ "Subtype", 
-    age ~ "Age",
+    subtype ~ "Subtype",
+    age ~ "Age at diagnosis",
     race ~ "Race",
-    marriage_status ~ "Marital Status", 
-    n_stage ~ "N Stage", 
+    marriage_status ~ "Marital Status",
+    n_stage ~ "N Stage",
     m_stage ~ "M Stage",
     grade ~ "Grade",
-    radiation ~ "Radiation", 
+    radiation ~ "Radiation",
     chemotherapy ~ "Chemotherapy",
-    surgery ~ "Surgery"
-  ),
-  hide_n = TRUE,
-)
+    surgery ~ "Surgery"),
+  statistic = "{estimate}",
+  times = c(0, 12, 60, 120)
+  ) |>
+  add_p()
 
 # create and assign more descriptive labels
 labels = c(
@@ -39,51 +43,44 @@ labels = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-t3_uv$table_body$label[1:36] <- labels
+t3$table_body$label[1:36] <- labels
+t3$table_styling$header$label[6:9] <- c("**Median Survival (years)**", "**1 Year**", "**5 Year**", "**10 Year**")
 
-# next multivariable table
-t3_mv <- coxph(
-  Surv(survival_months, survival_status) ~ 
-    subtype + 
-    age + 
-    race + 
-    marriage_status + 
-    n_stage + 
-    m_stage + 
[…]

### 2024-12-02 code/analysis/table4.R

@@ -1,0 +1,90 @@
+# Table 4: Cox Proportional Hazards Model Regression Results
+
+# load data
+seer <- readRDS("data/processed/seer.rds")
+attach(seer)
+
+# first univariate table
+t4_uv <- gtsummary::tbl_uvregression(
+  data = seer,
+  method = coxph,
+  y = Surv(survival_months, survival_status),
+  exponentiate = TRUE,
+  include = c(subtype, age, race, marriage_status, n_stage, m_stage, grade, radiation, chemotherapy, surgery),
+  label = list(
+    subtype ~ "Subtype", 
+    age ~ "Age",
+    race ~ "Race",
+    marriage_status ~ "Marital Status", 
+    n_stage ~ "N Stage", 
+    m_stage ~ "M Stage",
+    grade ~ "Grade",
+    radiation ~ "Radiation", 
+    chemotherapy ~ "Chemotherapy",
+    surgery ~ "Surgery"
+  ),
+  hide_n = TRUE,
+)
+
+# create and assign more descriptive labels
+labels = c(
+  "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
+  "Age at diagnosis", "20-49 years", "50-69 years", ">=70 years",
+  "Race", "White", "Black", "Other",
+  "Marital Status", "Unmarried", "Married",
+  "N Stage", "N0", "N1", "N2", "N3",
+  "M Stage", "M0", "M1",
+  "Grade", "I/II", "III/IV",
+  "Radiation", "No/Unknown", "Yes",
+  "Chemotherapy", "No/Unknown", "Yes",
+  "Surgery", "No/Unknown", "Yes")
+
+t4_uv$table_body$label[1:36] <- labels
+
+# next multivariable table
+t4_mv <- coxph(
+  Surv(survival_months, survival_status) ~ 
+    subtype + 
+    age + 
+    race + 
+    marriage_status + 
+    n_stage + 
+    m_stage + 
+    grade + 
+    radiation + 
+    chemotherapy + 
+    surgery,
+  data = seer, 
+  method = "breslow") |>
+  gtsummary::tbl_regression(
+    exp = TRUE,
+    label = list(
+      subtype ~ "Subtype", 
+      age ~ "Age",
+      race ~ "Race",
+      marriage_status ~ "Marital Status", 
+      n_stage ~ "N Stage", 
+      m_stage ~ "M Stage",
[…]

### 2024-12-03 products/manuscript/manuscript.qmd
manuscript edits

@@
 ---
-title: "epid 8010 manuscript"
-author: " "
+title: "Impact of Molecular Subtype on Inflammatory Breast Cancer-Specific Mortality in Adult Women"
+author: "Murphy John"
 format: pdf
@@
 ```
-
-**Author**
-
-Murphy John$^{1}$
-
-**Author affiliation**
-
-1.  Department of Epidemiology and Biostatistics, University of Georgia, Athens, GA, USA.
 
@@
 
-Therefore, using the most recent Surveillance, Epidemiology, and End Results (SEER) data, this study aims to assess the effect of IBC molecular subtype on breast cancer-specific mortality in adult females with IBC.
+Therefore, using the most recent Surveillance, Epidemiology, and End Results (SEER) data, this study aims to assess the effect of molecular subtype on IBC-specific mortality in adult females with IBC.
 
@@
 
+\clearpage
+
 # Results
@@
 Of these, 1829 were excluded; 165 had a survival time less than 1 month, 487 had an unknown or competing cause of death, and 1177 had incomplete information.
-The remaining 4140 were included in our analysis, as shown in Figure 1.
+The remaining 4140 were included in our analysis, as shown in @fig-flowchart.
 
@@
 #|   [Flowchart of IBC patients screening on SEER database]{color="#FF2400"}
-#| out.width: 100%
-#| out.height: 80%
+#| out.width: 90%
+#| out.height: 70%
 knitr::include_graphics(
@@
 
-The clinicopathological characteristics of eligible patients are summarized in Table 1.
+The clinicopathological characteristics of eligible patients are summarized in @tbl-table1.
 The age subgroup of 20–49 years accounts for 35.1% of Luminal B cases, a higher proportion compared to other subtypes.
@@
 #|   [Clinicopathological characteristics by cancer subtype]{color="#FF2400"}
+#| echo: false
+
 knitr::include_graphics(
-	here::here("results", "tables", "t1.png")
+	here::here("results", "tables", "t1.rds")
 )
@@
 
+\clearpage
+
 ## Proportional Hazards Assumption
 
-The proportional hazards assumption was assessed for each covariate visually using log minus log plots, as shown in Figure 2.
+The proportional hazards assumption was assessed for each covariate visually using log minus log plots, as shown in @fig-loglog.
 
 ```{r}
-#| label: fig-flowchart
+#| label: fig-loglog
 #| fig-cap: |
-#|   [Log-log plots]{color="#FF2400"}
+#|   [Log minus log plots]{color="#FF2400"}
 #| out.width: 100%
-#| out.height: 80%
 knitr::include_graphics(
@@
 
-Upon inspection, slight deviations from parallelism were observed for the covariates subtype, radiation, chemotherapy, and surgery.
+Upon inspection, slight deviations from parallelism were observed for the covariates subtype, grade radiation, chemotherapy, and surgery.
 Therefore, the full Cox proportional hazards model was fitted to the data, and Schoenfeld residuals were extracted.
-The correlation between Schoenfeld residuals and ranked survival times for each covariate is presented in Table 2.
+The correlation between Schoenfeld residuals and ranked survival times for each covariate is presented in @tbl-table2.
+
+```{r}
+#| label: tbl-table2
+#| tbl-cap: |
+#|   [Schoenfeld Residual and Ranked Survival Time Correlation]{color="#FF2400"}
+#| out.width: 40%
+knitr::include_graphics(
+	here::here("results", "tables", "t2.png")
+)
+```
+
+While the p-values for the covariates triple negative subtype, grade, radiation, chemotherapy, and surgery are statistically significant at the 0.05 level, the observed correlations are not of large magnitude.
+Specifically, the correlations are -0.128, -0.042, 0.070, 0.110, and 0.074, respectively.
+Given that these values are relatively small in magnitude, we proceed with the Cox proportional hazards regression model.
+
+\clearpage
 
@@
 
-Table 2 presents the median survival in years, along with 1-, 5-, and 10-year survival percentages, stratified by clinicopathological characteristics, all of which show statistically significant differences based on the log-rank test.
+@tbl-table3 presents the median survival in years, along with 1-, 5-, and 10-year survival percentages, stratified by clinicopathological characteristics, all of which show statistically significant differences based on the log-rank test.
 Survival outcomes differ significantly by subtype, with the Triple-Negative subtype showing the poorest prognosis, including a median survival of 1.7 years and 1-, 5-, and 10-year survival rates of 67%, 24%, and 21%, respectively.
@@
 ```{r}
-#| label: tbl-table2
+#| label: tbl-table3
 #| tbl-cap: |
 #|   [Survival for IBC by Clinicopathological Characteristics]{color="#FF2400"}
+#| out.width: 100%
 knitr::include_graphics(
-	here::here("results", "tables", "t2.png")
+	here::here("results", "tables", "t3.png")
 )
@@
 
-We conducted univariate and multivariate analyses based on the data in Table 2, with the results of the Cox proportional hazards regression presented in Table 3.
+We conducted univariate and multivariate analyses based on the data in @tbl-table3, with the results of the Cox proportional hazards regression presented in @tbl-table4.
 
 ```{r}
-#| label: tbl-table3
+#| label: tbl-table4
 #| tbl-cap: |
@@
 #|   Model for Survival in Adult Women with IBC]{color="#FF2400"}
+#| out.width: 100%
 knitr::include_graphics(
-	here::here("results", "tables", "t3.png")
+	here::here("results", "tables", "t4.png")
 )
@@
 
+\clearpage
+
 # Discussion
@@
 In this large population-based cohort of women diagnosed with IBC, we found a better survival in Luminal B molecular subtype and poorer in Triple Negative.
-We found that Found that Black patients made up the smallest percentage of Luminal B subtype and the greatest of Triple Negative.
+We found that Black patients made up the smallest percentage of Luminal B subtype and the greatest of Triple Negative.
 Similar findings were observed for patients who were unmarried at diagnosis.
 
+Take things that ive noted from other study and compare results here.
 It would be interesting to look at the combination of treatments (radiation, chemo, surgery)
+Also comparing subtype to broader range of breat cancer not just IBC.

### 2024-12-03 products/manuscript/manuscript.tex
manuscript edits

@@
 \hypersetup{
-  pdftitle={epid 8010 manuscript},
+  pdftitle={Impact of Molecular Subtype on Inflammatory Breast Cancer-Specific Mortality in Adult Women},
+  pdfauthor={Murphy John},
   colorlinks=true,
@@
 
-\title{epid 8010 manuscript}
-\author{}
+\title{Impact of Molecular Subtype on Inflammatory Breast
+Cancer-Specific Mortality in Adult Women}
+\author{Murphy John}
 \date{}
@@
 \maketitle
-
-\textbf{Authors}
-
-\begin{itemize}
-\tightlist
-\item
-  Murphy John\(^{1}\)
-\end{itemize}
-
-\textbf{Author affiliation}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Department of Epidemiology and Biostatistics, University of Georgia,
-  Athens, GA, USA.
-\end{enumerate}
 
@@
 Therefore, using the most recent Surveillance, Epidemiology, and End
-Results (SEER) data, this study aims to assess the effect of IBC
-molecular subtype on breast cancer-specific mortality in adult females
-with IBC.
+Results (SEER) data, this study aims to assess the effect of molecular
+subtype on IBC-specific mortality in adult females with IBC.
 
@@
 For each case, the following information from SEER was obtained: age at
-diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other),
-marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage
-(M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes),
-chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival
-months, cause-specific death classification (Alive or dead of other
-cause, Dead (attributable to this cancer), Dead (missing/unknown COD)),
-and other cause of death classification (Alive or dead due to cancer,
-Dead (attributable to causes other than this cancer), Dead
-(missing/unknown COD)). Breast cancer subtypes were classified as
-Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive (HR-/HER2+),
-and Triple Negative. In addition, cases with incomplete information on
-any of these characteristics and patients who survived less than one
-month were excluded from the study. Additionally, to focus on breast
-cancer-specific mortality, patients with an unknown cause of death or
-whose death was attributed to other causes were excluded.
+diagnosis (20-49, 50-69, and \textgreater=70), race (white, black,
+other), marital status (unmarried, married), N (node) stage (N0, N1, N2,
+N3), M (metastasis) stage (M0, M1), grade (I/II, III/IV), radiation
+(no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown,
+yes), survival months, cause-specific death classification (Alive or
+dead of other cause, Dead (attributable to this cancer), Dead
+(missing/unknown COD)), and other cause of death classification (Alive
+or dead due to cancer, Dead (attributable to causes other than this
+cancer), Dead (missing/unknown COD)). Breast cancer subtypes were
+classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive
+(HR-/HER2+), and Triple Negative. In addition, cases with incomplete
+information on any of these characteristics and patients who survived
+less than one month were excluded from the study. To focus on
+IBC-specific mortality, patients with an unknown cause of death or whose
+death was attributed to other causes were excluded.
 
@@
 
-Clinicopatholical characteristics were compared between IBC subtype
+Clinicopathological characteristics were compared between IBC subtype
 using Pearson's Chi-squared test. Median, 1-year, 5-year, and 10-year
-survival for IBC is estimated for each variable in the analysis using
-the Kaplain-Meier method and variables were compared using the log-rank
-test in the subgroups. Univariate and multivariate Cox proportional
-hazards regressions were used to model the relationship between
-potential covariates and IBC-specific survival. We obtained hazard
-ratios (HR) and 95\% confidence intervals for each covariate. All
+survival for IBC is estimated for each variable in using the
+Kaplain-Meier method and compared using the log-rank test in the
+subgroups. The proportional hazards assumption was assessed visually
+with log-log plots and statistically by calculating Schoenfeld residuals
+and the Pearson's product moment correlation coefficient of the
+Schoenfeld residuals and ranked survival times. After determining the
+assumption of proportional hazard was met, univariate and multivariate
+Cox proportional hazards regressions were used to model the relationship
+between potential covariates and IBC-specific survival. We obtained
+hazard ratios (HR) and 95\% confidence intervals for each covariate. All
 statistical analyses were performed using the survival package in R
 version 4.3.3.
+
+\clearpage
 
@@
 or competing cause of death, and 1177 had incomplete information. The
-remaining 4140 were included in our analysis, as shown in Table 1. The
-clinicopathological characteristics of the eligible patients are shown
-in Table 2.
+remaining 4140 were included in our analysis, as shown in
+Figure~\ref{fig-flowchart}.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=0.9\textwidth,height=0.7\textheight]{../../results/figures/flowchart.png}
+
+}
+
+\caption{\label{fig-flowchart}{Flowchart of IBC patients screening on
+SEER database}}
+
+\end{figure}%
+
+The clinicopathological characteristics of eligible patients are
+summarized in Table~\ref{tbl-table1}. The age subgroup of 20--49 years
+accounts for 35.1\% of Luminal B cases, a higher proportion compared to
+other subtypes. Moreover, Luminal B and HER2-positive subtypes appear to
+affect younger age groups more frequently than Luminal A, with 85.4\%
+and 84.5\% of cases, respectively, occurring in patients under 69 years,
+compared to 76.5\% for Luminal A. A higher proportion of Black patients
+is observed in the Triple-Negative group compared to other subtypes.
+Marital status analysis shows that the Triple-Negative group has the
+highest percentage of unmarried women. Triple-Negative cases also have
+the highest percentage of N3 node stage and exhibit the greatest
+proportion of grade III/IV cases.
+
+\begin{table}
+
+\caption{\label{tbl-table1}{Clinicopathological characteristics by
+cancer subtype}}
+
+\centering{
+
+\includegraphics{../../results/tables/t1.rds}
+
+}
+
+\end{table}%
+
+\clearpage
+
+\subsection{Proportional Hazards
+Assumption}\label{proportional-hazards-assumption}
+
+The proportional hazards assumption was assessed for each covariate
+visually using log minus log plots, as shown in Figure~\ref{fig-loglog}.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{../../results/figures/loglog_plot.png}
+
+}
+
+\caption{\label{fig-loglog}{Log minus log plots}}
+
+\end{figure}%
+
+Upon inspection, slight deviations from parallelism were observed for
+the covariates subtype, grade radiation, chemotherapy, and surgery.
+Therefore, the full Cox proportional hazards model was fitted to the
+data, and Schoenfeld residuals were extracted. The correlation between
+Schoenfeld residuals and ranked survival times for each covariate is
+presented in Table~\ref{tbl-table2}.
+
+\begin{table}
+
+\caption{\label{tbl-table2}{Schoenfeld Residual and Ranked Survival Time
+Correlation}}
+
+\centering{
+
+\includegraphics[width=0.4\textwidth,height=\textheight]{../../results/tables/t2.png}
+
+}
+
+\end{table}%
+
+While the p-values for the covariates triple negative subtype, grade,
+radiation, chemotherapy, and surgery are statistically significant at
+the 0.05 level, the observed correlations are not of large magnitude.
+Specifically, the correlations are -0.128, -0.042, 0.070, 0.110, and
+0.074, respectively. Given that these values are relatively small in
+magnitude, we proceed with the Cox proportional hazards regression
+model.
+
+\clearpage
 
@@
 
+Table~\ref{tbl-table3} presents the median survival in years, along with
+1-, 5-, and 10-year survival percentages, stratified by
+clinicopathological characteristics, all of which show statistically
+significant differences based on the log-rank test. Survival outcomes
+differ significantly by subtype, with the Triple-Negative subtype
+showing the poorest prognosis, including a median survival of 1.7 years
+and 1-, 5-, and 10-year survival rates of 67\%, 24\%, and 21\%,
+respectively. Conversely, the Luminal B subtype has the most favorable
+outcomes, with a median survival of 9.4 years and a 10-year survival
+rate of 50\%. Racial disparities are evident, as Black patients have the
+poorest survival outcomes, with a median survival of 2.4 years compared
+to 4.6 years for White patients and 5.4 years for patients of other
+races. Unmarried women experience poorer survival compared to married
+women. Although 1-year survival is similar across all node stages, nodal
+involvement significantly affects long-term outcomes, with the N3 stage
+showing the lowest median survival of 2.7 years and the poorest 5- and
+10-year survival rates. Survival differences are also pronounced by
+distant metastasis (M stage), with M1 stage associated with worse
+outcomes. Additionally, higher cancer grades correlate with poorer
+survival. In terms of treatment, patients who receive radiation,
+chemotherapy, or surgery exhibit better survival outcomes across all
+measurements.
+
+\begin{table}
+
+\caption{\label{tbl-table3}{Survival for IBC by Clinicopathological
+Characteristics}}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{../../results/tables/t3.png}
+
+}
+
+\end{table}%
+
+We conducted univariate and multivariate analyses based on the data in
+Table~\ref{tbl-table3}, with the results of the Cox proportional hazards
+regression presented in Table~\ref{tbl-table4}.
+
+\begin{table}
+
+\caption{\label{tbl-table4}{Cox Proportional Hazards Regression Model
+for Survival in Adult Women with IBC}}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{../../results/tables/t4.png}
+
+}
+
+\end{table}%
+
 \subsection{Univariate Analysis}\label{univariate-analysis}
+
+In the univariate analysis, subtype, age, race, marital status, N stage,
+M stage, grade, radiation, chemotherapy, and surgery were all
+significantly associated with IBC-specific survival. These variables
+were therefore included in the multivariate analysis.
 
@@
 
-\section*{Discussion}\label{discussion}
-\addcontentsline{toc}{section}{Discussion}
+The multivariate analysis revealed that subtype, age, race, marital
+status, N stage, M stage, grade, radiation, chemotherapy, and surgery
+were independent predictors of IBC-specific survival. The Luminal B
+subtype was associated with better survival (aHR=0.72, 95\%CI:
+0.56-0.74, P\textless0.0001), whereas the Triple-Negative subtype showed
+the poorest survival outcomes (aHR=2.51, 95\%CI: 1.79-2.21,
+P\textless0.0001). In terms of treatment, our results find that the
+existing therapies are effective in improving survival, that is
+radiation (aHR=0.76, 95\%CI: 0.69-0.85, P\textless0.0001), chemotherapy
+(aHR=0.50, 95\%CI: 0.44-0.57, P\textless0.0001), and surgery (aHR=0.59,
+95\%CI: 0.52-0.66, P\textless0.0001).
+
+\clearpage
+
+\section{Discussion}\label{discussion}
+
+In this large population-based cohort of women diagnosed with IBC, we
+found a better survival in Luminal B molecular subtype and poorer in
+Triple Negative. We found that Black patients made up the smallest
+percentage of Luminal B subtype and the greatest of Triple Negative.
+Similar findings were observed for patients who were unmarried at
+diagnosis.
+
+Take things that ive noted from other study and compare results here. It
+would be interesting to look at the combination of treatments
+(radiation, chemo, surgery) Also comparing subtype to broader range of
+breat cancer not just IBC.
 

### 2024-12-03 code/analysis/table1.R

@@ -13,7 +13,8 @@ box::use(
   webshot2,
   tidyr,
   ggplot2,
-  Hmisc
+  Hmisc,
+  flextable
 )
 
 # load data
@@ -61,3 +62,7 @@ gt::gtsave(
   vwidth = 2400,
   vheight = 1350
 )
+
+# save to rds
+t1_flex <- flextable::qflextable(t1)
+saveRDS(t1_flex, file = "results/tables/t1.rds")

### 2024-12-03 code/analysis/table2.R

@@ -3,6 +3,7 @@
 library(survival)
 library(dplyr)
 library(gtsummary)
+library(gt)
 
 # load data
 seer <- readRDS("data/processed/seer.rds")
@@ -14,7 +15,7 @@ model <- coxph(
     relevel(factor(subtype), ref = '0') +
     relevel(factor(age), ref = '0') + 
     relevel(factor(race), ref = '0') +
-    relevel(factor(marriage_status),ref='9') + 
+    relevel(factor(marriage_status),ref='0') + 
     relevel(factor(n_stage),ref='0') + 
     relevel(factor(m_stage),ref='0') +
     relevel(factor(grade), ref='0') +
@@ -27,21 +28,30 @@ model <- coxph(
 # extract Schoenfeld residuals
 resid <- cox.zph(model, terms = FALSE)
 
-# get the correlations between residuals and ranked survival times
-cor <- data.frame(apply(resid$y, 2, \(x) cor.test(rank(resid$time), x)$estimate))
-
 # clean up to present in table
-rownames(cor) <- c("Luminal B", "Triple Negative", "HER2 Positive", 
-               "50-69 years", ">=70 years",
-               "Black", "Other",
-               "Married",
-               "N1", "N2", "N3",
-               "M1",
-               "III/IV",
-               "Radiation: Yes",
-               "Chemotherapy: Yes",
-               "Surgery: Yes")
-colnames(cor) <- "correlation"  
+covariates <- c("Luminal B", "Triple Negative", "HER2 Positive", 
+                "50-69 years", ">=70 years",
+                "Black", "Other",
+                "Married",
+                "N1", "N2", "N3",
+                "M1",
+                "Grade: III/IV",
+                "Radiation: Yes",
+                "Chemotherapy: Yes",
+                "Surgery: Yes")
 
-t2 <- gtsummary::tbl_summary(cor)
+# get the correlations between residuals and ranked survival times
+cor <- data.frame(
+  Variable = covariates,
+  Correlation = apply(resid$y, 2, \(x) round(cor.test(rank(resid$time), x)$estimate, 3)),
+  p.value = apply(resid$y, 2, \(x) {
+    p <- round(cor.test(rank(resid$time), x)$p.value, 3)
+    ifelse(p == 0, "<0.001", p)
+  })
+)
+
+t2 <- cor %>% gt()
+
+# save as png
+gtsave(data = t2, filename = "results/tables/t2.png")
 

### 2024-12-03 products/manuscript/manuscript.qmd
tibbles 1 and 2

@@
 library(here)
+library(flextable)
 
@@
 
-\clearpage
+\pagebreak
 
@@
 #| echo: false
+#| results: asis
 
-knitr::include_graphics(
-	here::here("results", "tables", "t1.rds")
-)
+t1 <- readRDS(file = here::here("results/tables", "t1.rds"))
+t1
+
 ```
-
-\clearpage
 
@@
 
+\clearpage
+
 Upon inspection, slight deviations from parallelism were observed for the covariates subtype, grade radiation, chemotherapy, and surgery.
@@
 #|   [Schoenfeld Residual and Ranked Survival Time Correlation]{color="#FF2400"}
-#| out.width: 40%
-knitr::include_graphics(
-	here::here("results", "tables", "t2.png")
-)
+#| echo: false
+#| results: asis
+
+t2 <- readRDS(file = here::here("results/tables", "t2.rds"))
+t2
 ```
@@
 Given that these values are relatively small in magnitude, we proceed with the Cox proportional hazards regression model.
-
-\clearpage
 

### 2024-12-03 code/analysis/table1.R

@@ -2,34 +2,27 @@
 # (unadjusted associations between covariates and exposure)
 
 # Declare dependencies
-box::use(
-  readr,
-  here,
-  gt,
-  gtExtras,
-  gtsummary,
-  modelsummary,
-  dplyr,
-  webshot2,
-  tidyr,
-  ggplot2,
-  Hmisc,
-  flextable
-)
+library(here)
+library(readr)
+library(gt)
+library(gtsummary)
+library(gtExtras)
+library(flextable)
+library(stringr)
 
 # load data
 seer <- readRDS("data/processed/seer_tab.rds")
 attach(seer)
 
 # create table 1
-t1 <-
+t1 <- gtsummary::as_tibble(
   seer |>
   gtsummary::tbl_summary(
     by = subtype,
     include = c(agecat, race, marriage_status, n_stage, m_stage,
                 gradecat, radiation, chemotherapy, surgery),
     label = list(
-      agecat ~ "Age at diagnosis",
+      agecat ~ "Age",
       race ~ "Race",
       marriage_status ~ "Marital Status",
       n_stage ~ "N Stage",
@@ -52,17 +45,14 @@ t1 <-
                gradecat, radiation, chemotherapy, surgery) ~ c(0, 1),
   ) |>
   gtsummary::add_overall() |>
-  gtsummary::add_p() |>
-  gtsummary::as_gt()
-
-# save to png
-gt::gtsave(
-  data = t1,
-  filename = here::here("results/tables", "t1.png"),
-  vwidth = 2400,
-  vheight = 1350
+  gtsummary::add_p()
 )
 
-# save to rds
-t1_flex <- flextable::qflextable(t1)
-saveRDS(t1_flex, file = "results/tables/t1.rds")
+# Remove Markdown bold (`**`) from the tibble text
+colnames(t1) <- stringr::str_remove_all(colnames(t1), "\\*\\*")
+
+# Convert the data frame to a flextable
+t1_flex <- flextable::flextable(t1)
+t1_flex <- flextable::fontsize(t1_flex, size = 7.5, part = "all")
+# Save the flextable object as RDS
+saveRDS(t1_flex, file = here::here("results/tables", "t1.rds"))

### 2024-12-03 code/analysis/table2.R

@@ -44,14 +44,17 @@ covariates <- c("Luminal B", "Triple Negative", "HER2 Positive",
 cor <- data.frame(
   Variable = covariates,
   Correlation = apply(resid$y, 2, \(x) round(cor.test(rank(resid$time), x)$estimate, 3)),
-  p.value = apply(resid$y, 2, \(x) {
+  `p-value` = apply(resid$y, 2, \(x) {
     p <- round(cor.test(rank(resid$time), x)$p.value, 3)
     ifelse(p == 0, "<0.001", p)
   })
 )
 
-t2 <- cor %>% gt()
+t2 <- gtsummary::as_tibble(cor)
 
-# save as png
-gtsave(data = t2, filename = "results/tables/t2.png")
+# Convert to a flextable
+t2_flex <- flextable::flextable(t2)
+t2_flex <- flextable::fontsize(t2_flex, size = 7.5, part = "all")
+# Save the flextable object as RDS
+saveRDS(t2_flex, file = here::here("results/tables", "t2.rds"))
 

### 2024-12-04 products/manuscript/manuscript.qmd
tibbles 3 and 4

@@
 ---
-title: "Impact of Molecular Subtype on Inflammatory Breast Cancer-Specific Mortality in Adult Women"
+title: "Impact of Molecular Subtype on Inflammatory Breast Cancer Mortality in Adult Women"
 author: "Murphy John"
@@
 In terms of treatment, patients who receive radiation, chemotherapy, or surgery exhibit better survival outcomes across all measurements.
-
 ```{r}
@@
 #|   [Survival for IBC by Clinicopathological Characteristics]{color="#FF2400"}
-#| out.width: 100%
-knitr::include_graphics(
-	here::here("results", "tables", "t3.png")
-)
+#| echo: false
+#| results: asis
+
+t3 <- readRDS(file = here::here("results/tables", "t3.rds"))
+t3
 ```

### 2024-12-04 code/analysis/table3.R

@@ -9,14 +9,15 @@ seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
 # 1, 5, and 10 year survival
-t3 <- tbl_survfit(
-  seer,
+t3 <- gtsummary::as_tibble(
+  seer |>
+  gtsummary::tbl_survfit(
   y = "Surv(survival_months, survival_status)",
   include = c(subtype, age, race, marriage_status, n_stage, m_stage,
               grade, radiation, chemotherapy, surgery),
   label = list(
     subtype ~ "Subtype",
-    age ~ "Age at diagnosis",
+    age ~ "Age",
     race ~ "Race",
     marriage_status ~ "Marital Status",
     n_stage ~ "N Stage",
@@ -29,11 +30,14 @@ t3 <- tbl_survfit(
   times = c(0, 12, 60, 120)
   ) |>
   add_p()
+)
+
+colnames(t3) <- c("Characteristic","Median Survival (years)", "1 Year", "5 Year", "10 Year", "p-value")
 
 # create and assign more descriptive labels
-labels = c(
+t3$Characteristic = c(
   "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
-  "Age at diagnosis", "20-49 years", "50-69 years", ">=70 years",
+  "Age", "20-49 years", "50-69 years", ">=70 years",
   "Race", "White", "Black", "Other",
   "Marital Status", "Unmarried", "Married",
   "N Stage", "N0", "N1", "N2", "N3",
@@ -43,45 +47,42 @@ labels = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-t3$table_body$label[1:36] <- labels
-t3$table_styling$header$label[6:9] <- c("**Median Survival (years)**", "**1 Year**", "**5 Year**", "**10 Year**")
-
 # since we can't do both median survival and specific times,
 # create a seperate table with median survival
-t3_med <- tbl_survfit(
-  seer,
-  y = "Surv(survival_months, survival_status)",
-  include = c(subtype, age, race, marriage_status, n_stage, m_stage,
-              grade, radiation, chemotherapy, surgery),
-  label = list(
-    subtype ~ "Subtype",
-    age ~ "Age at diagnosis",
-    race ~ "Race",
-    marriage_status ~ "Marital Status",
-    n_stage ~ "N Stage",
-    m_stage ~ "M Stage",
-    grade ~ "Grade",
[…]

### 2024-12-04 code/analysis/table4.R

@@ -24,12 +24,14 @@ t4_uv <- gtsummary::tbl_uvregression(
     surgery ~ "Surgery"
   ),
   hide_n = TRUE,
-)
+  )
+
+#colnames(t4_uv) <- c("Characteristic","HR", "95% CI", "p-value")
 
 # create and assign more descriptive labels
-labels = c(
+t4_uv$table_body$label[1:36] = c(
   "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
-  "Age at diagnosis", "20-49 years", "50-69 years", ">=70 years",
+  "Age", "20-49 years", "50-69 years", ">=70 years",
   "Race", "White", "Black", "Other",
   "Marital Status", "Unmarried", "Married",
   "N Stage", "N0", "N1", "N2", "N3",
@@ -39,52 +41,100 @@ labels = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-t4_uv$table_body$label[1:36] <- labels
-
 # next multivariable table
 t4_mv <- coxph(
-  Surv(survival_months, survival_status) ~ 
-    subtype + 
-    age + 
-    race + 
-    marriage_status + 
-    n_stage + 
-    m_stage + 
-    grade + 
-    radiation + 
-    chemotherapy + 
-    surgery,
-  data = seer, 
-  method = "breslow") |>
-  gtsummary::tbl_regression(
-    exp = TRUE,
-    label = list(
-      subtype ~ "Subtype", 
-      age ~ "Age",
-      race ~ "Race",
-      marriage_status ~ "Marital Status", 
-      n_stage ~ "N Stage", 
-      m_stage ~ "M Stage",
-      grade ~ "Grade",
-      radiation ~ "Radiation", 
-      chemotherapy ~ "Chemotherapy",
-      surgery ~ "Surgery"
+    Surv(survival_months, survival_status) ~ 
+      subtype + 
+      age + 
+      race + 
+      marriage_status + 
+      n_stage + 
+      m_stage + 
+      grade + 
+      radiation + 
+      chemotherapy + 
+      surgery,
+    data = seer, 
+    method = "breslow") |>
+    gtsummary::tbl_regression(
+      exp = TRUE,
+      label = list(
+        subtype ~ "Subtype", 
+        age ~ "Age",
+        race ~ "Race",
+        marriage_status ~ "Marital Status", 
+        n_stage ~ "N Stage", 
+        m_stage ~ "M Stage",
[…]

### 2024-12-05 products/manuscript/manuscript.qmd
first draft done

@@
 ---
-title: "Impact of Molecular Subtype on Inflammatory Breast Cancer Mortality in Adult Women"
+title: |
+  \LARGE Effect of Molecular Subtype on Mortality in U.S. Adult Women with Inflammatory Breast Cancer
+subtitle: "EPID 8010 Final Project Fall 2024"
 author: "Murphy John"
@@
   - "refs.bib"
+csl: "ieee.csl"
+nocite: "@*"
 ---
@@
 
-\pagebreak
-
 # Abstract
@@
 **Background:** 
+Inflammatory breast cancer (IBC) is a rare and aggressive form of breast cancer characterized by rapid progression and poor prognosis.
+Molecular subtypes of breast cancer, such as Luminal A, Luminal B, HER2 Positive, and Triple Negative, have been associated with varying survival outcomes.
+However, the impact of these subtypes specifically on IBC mortality remains underexplored.
 
 **Methods:** 
+We conducted a retrospective cohort study of patients diagnosed with IBC using the SEER database.
+The cohort was stratified by molecular subtype, age, race, marital status, nodal stage, distant metastasis, and treatment modalities.
+Cox proportional hazards regression models were used to assess the association between molecular subtype and IBC-specific survival, adjusting for potential confounders.
 
 **Results:**
+Our findings demonstrate significant survival differences across molecular subtypes.
+The Triple Negative subtype exhibited the poorest prognosis (aHR=2.51, 95%CI: 1.79-2.21, P\<0.001), with a median survival of 1.7 years and 1-, 5-, and 10-year survival rates of 67%, 24%, and 21%, respectively.
+In contrast, the Luminal B subtype had the best outcomes (aHR=0.72, 95%CI: 0.56-0.74, P\<0.001), with a median survival of 9.4 years and a 10-year survival rate of 50%.
 
 **Conclusions:**
+These findings highlight the importance of subtype classification in prognostication and treatment planning for IBC.
 
@@
 Studies have shown that these four molecular subtypes reveal the prognostic discrepancies in both common breast cancers and IBC [@robertson2010inflammatory; @le2015future; @tolaney2015adjuvant; @iwamoto2011different].
-A previous study examined the association between molecular subtypes and survival outcomes in IBC patients from 2010 to 2013 [@li2017outcomes].
-However, no updated population-based studies have investigated this relationship.
+A previous study examined the association between molecular subtype and survival outcomes in IBC patients using SEER data from 2010 to 2013 [@li2017outcomes].
+However, no studies utilizing the updated SEER database have investigated this relationship.
 
@@
 Year of diagnosis was limited to 2010 to 2021 because SEER started collecting breast cancer molecular subtype in 2010.
-To limit our data to adult females, age at diagnosis was limited to patents 20 years and older and sex was limited to "Female".
+To limit our data to adult females, age at diagnosis was limited to patients 20 years and older and sex was limited to "Female".
 
@@
 
-For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N (node) stage (N0, N1, N2, N3), M (metastasis) stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer), Dead (missing/unknown COD)).
-Breast cancer subtypes were classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 positive (HR-/HER2+), and Triple Negative.
+For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N (node) stage (N0, N1, N2, N3), M (metastasis) stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), and survival months.
+Breast cancer subtypes were classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 Positive (HR-/HER2+), and Triple Negative.
 In addition, cases with incomplete information on any of these characteristics and patients who survived less than one month were excluded from the study.
@@
 Clinicopathological characteristics were compared between IBC subtype using Pearson's Chi-squared test.
-Median, 1-year, 5-year, and 10-year survival for IBC is estimated for each variable in using the Kaplain-Meier method and compared using the log-rank test in the subgroups.
-The proportional hazards assumption was assessed visually with log-log plots and statistically by calculating Schoenfeld residuals and the Pearson's product moment correlation coefficient of the Schoenfeld residuals and ranked survival times.
-After determining the assumption of proportional hazard was met, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
-We obtained hazard ratios (HR) and 95% confidence intervals for each covariate.
-All statistical analyses were performed using the survival package in R version 4.3.3.
+Median, 1-, 5-, and 10-year survival for IBC is estimated for each variable using the Kaplain-Meier method and compared using the log-rank test in the subgroups.
+The proportional hazards assumption was assessed visually with log minus log plots and statistically by calculating the Pearson's product moment correlation coefficient of the Schoenfeld residuals and ranked survival times.
+After determining the assumption of proportional hazard was not violated, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
+We obtained hazard ratios (HR) and 95% confidence intervals (CI) for each covariate.
+All statistical analyses were performed using version 3.7-0 of the survival package in R version 4.3.3.
 
@@
 
-A total of 5969 adult female IBC cases were eligible during 2010-2021 in the SEER database for our study population.
+A total of 5969 adult female IBC cases were eligible from 2010-2021 in the SEER database for our study population.
 Of these, 1829 were excluded; 165 had a survival time less than 1 month, 487 had an unknown or competing cause of death, and 1177 had incomplete information.
@@
 The age subgroup of 20–49 years accounts for 35.1% of Luminal B cases, a higher proportion compared to other subtypes.
-Moreover, Luminal B and HER2-positive subtypes appear to affect younger age groups more frequently than Luminal A, with 85.4% and 84.5% of cases, respectively, occurring in patients under 69 years, compared to 76.5% for Luminal A.
+Moreover, Luminal B and HER2 Positive subtypes appear to affect younger age groups more frequently than Luminal A, with 85.4% and 84.5% of cases, respectively, occurring in patients under 69 years, compared to 76.5% for Luminal A.
 A higher proportion of Black patients is observed in the Triple-Negative group compared to other subtypes.
-Marital status analysis shows that the Triple-Negative group has the highest percentage of unmarried women.
-Triple-Negative cases also have the highest percentage of N3 node stage and exhibit the greatest proportion of grade III/IV cases.
+Comparing marital status reveals that the Triple Negative subtype affects the highest percentage of unmarried women.
+Triple Negative patients also have the highest proportion of N3 node stage and cancer grade III/IV.
 
@@
 
-Upon inspection, slight deviations from parallelism were observed for the covariates subtype, grade radiation, chemotherapy, and surgery.
-Therefore, the full Cox proportional hazards model was fitted to the data, and Schoenfeld residuals were extracted.
+Upon inspection, slight deviations from parallelism were observed for the covariates subtype, grade, radiation, chemotherapy, and surgery.
+Therefore, Schoenfeld residuals were extracted from the full Cox proportional hazards model.
 The correlation between Schoenfeld residuals and ranked survival times for each covariate is presented in @tbl-table2.
@@
 
-While the p-values for the covariates triple negative subtype, grade, radiation, chemotherapy, and surgery are statistically significant at the 0.05 level, the observed correlations are not of large magnitude.
-Specifically, the correlations are -0.128, -0.042, 0.070, 0.110, and 0.074, respectively.
+While the p-values for Triple Negative subtype, grade, radiation, chemotherapy, and surgery are statistically significant at the 0.05 level, the observed correlations are not of large magnitude.
+Respectively, the correlations are -0.128, -0.042, 0.070, 0.110, and 0.074.
 Given that these values are relatively small in magnitude, we proceed with the Cox proportional hazards regression model.
@@
 @tbl-table3 presents the median survival in years, along with 1-, 5-, and 10-year survival percentages, stratified by clinicopathological characteristics, all of which show statistically significant differences based on the log-rank test.
-Survival outcomes differ significantly by subtype, with the Triple-Negative subtype showing the poorest prognosis, including a median survival of 1.7 years and 1-, 5-, and 10-year survival rates of 67%, 24%, and 21%, respectively.
-Conversely, the Luminal B subtype has the most favorable outcomes, with a median survival of 9.4 years and a 10-year survival rate of 50%.
-Racial disparities are evident, as Black patients have the poorest survival outcomes, with a median survival of 2.4 years compared to 4.6 years for White patients and 5.4 years for patients of other races.
+Survival outcomes differ significantly by subtype, with the Triple Negative subtype showing the poorest prognosis, including a median survival of 1.7 years and 1-, 5-, and 10-year survival rates of 67%, 24%, and 21%, respectively.
+Conversely, the Luminal B subtype has the best outcomes, with a median survival of 9.4 years and a 10-year survival rate of 50%.
+Racial disparities are evident in the data, as Black patients have the poorest survival outcomes, with a median survival of 2.4 years compared to 4.6 years for White patients and 5.4 years for patients of other races.
 Unmarried women experience poorer survival compared to married women.
 Although 1-year survival is similar across all node stages, nodal involvement significantly affects long-term outcomes, with the N3 stage showing the lowest median survival of 2.7 years and the poorest 5- and 10-year survival rates.
-Survival differences are also pronounced by distant metastasis (M stage), with M1 stage associated with worse outcomes.
+Survival differences are also pronounced by distant metastasis, with M1 stage associated with worse outcomes.
 Additionally, higher cancer grades correlate with poorer survival.
@@
 
-We conducted univariate and multivariate analyses based on the data in @tbl-table3, with the results of the Cox proportional hazards regression presented in @tbl-table4.
+We conducted univariate and multivariate analyses based on the data in @tbl-table3. The results of the Cox proportional hazards regression models are presented in @tbl-table4.
 
@@
 #| tbl-cap: |
-#|   [Cox Proportional Hazards Regression
-#|   Model for Survival in Adult Women with IBC]{color="#FF2400"}
-#| out.width: 100%
-knitr::include_graphics(
-	here::here("results", "tables", "t4.png")
-)
+#|   [Crude and Adjusted Cox Proportional Hazards Regression
+#|   Models]{color="#FF2400"}
+#| echo: false
+#| results: asis
+
+t4 <- readRDS(file = here::here("results/tables", "t4.rds"))
+t4
 ```
@@
 The multivariate analysis revealed that subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were independent predictors of IBC-specific survival.
-The Luminal B subtype was associated with better survival (aHR=0.72, 95%CI: 0.56-0.74, P\<0.0001), whereas the Triple-Negative subtype showed the poorest survival outcomes (aHR=2.51, 95%CI: 1.79-2.21, P\<0.0001).
-In terms of treatment, our results find that the existing therapies are effective in improving survival, that is radiation (aHR=0.76, 95%CI: 0.69-0.85, P\<0.0001), chemotherapy (aHR=0.50, 95%CI: 0.44-0.57, P\<0.0001), and surgery (aHR=0.59, 95%CI: 0.52-0.66, P\<0.0001).
+The Luminal B subtype was associated with better survival (aHR=0.72, 95%CI: 0.56-0.74, P\<0.001), whereas the Triple Negative subtype showed the poorest survival outcomes (aHR=2.51, 95%CI: 1.79-2.21, P\<0.001).
+In terms of treatment, our results found that the existing therapies are effective in improving survival.
+That is, radiation (aHR=0.76, 95%CI: 0.69-0.85, P\<0.001), chemotherapy (aHR=0.50, 95%CI: 0.44-0.57, P\<0.001), and surgery (aHR=0.59, 95%CI: 0.52-0.66, P\<0.001).
 
@@
 
-In this large population-based cohort of women diagnosed with IBC, we found a better survival in Luminal B molecular subtype and poorer in Triple Negative.
-We found that Black patients made up the smallest percentage of Luminal B subtype and the greatest of Triple Negative.
-Similar findings were observed for patients who were unmarried at diagnosis.
+In this large population-based cohort of adult women diagnosed with IBC, we found improved survival in Luminal B molecular subtype and poorer in Triple Negative.
+These findings are consistent with previous literature, which similarly identified Triple Negative IBC as a particularly aggressive form of breast cancer with poorer prognoses compared to hormone receptor-positive subtypes [@li2017outcomes, @pan2019nomogram, @menta2018inflammatory].
 
-Take things that ive noted from other study and compare results here.
-It would be interesting to look at the combination of treatments (radiation, chemo, surgery)
-Also comparing subtype to broader range of breat cancer not just IBC.
+The racial disparities observed in our study, wherein Black participants had the poorest outcomes, align with existing evidence of inequities in IBC survival rates by race [@abraham2021incidence, @hance2005trends].
+This result might be due to the higher incidence of Triple Negative breast cancer observed in Black American women, a finding observed in this study and in prior works [@lund2009race].
+Furthermore, unmarried women experienced worse outcomes, consistent with studies suggesting that social support and marital status are important predictors of cancer survival [@osborne2005influence, @martinez2017prognostic].
+
+Treatment modalities, including radiation, chemotherapy, and surgery, were independently associated with improved survival outcomes in this study.
+These results are consistent with prior research demonstrating the efficacy of multimodal therapy in managing IBC [@rosso2017improved].
+However, the significantly poorer outcomes for the Triple Negative subtype highlight a need for improved therapeutic strategies targeting this group.
+
+# Limitations
+
+This study has several strengths, including the use of a large, diverse patient population and robust multivariate analyses that accounted for key clinicopathological factors.
+However, several limitations should be acknowledged.
+First, this study excluded 1177 individuals diagnosed with IBC due to incomplete data, which may introduce bias if the missing data were not random.
+Additionally, since this study focused on IBC-specific survival, individuals with unknown causes of death or those who died from causes unrelated to IBC were excluded (n=487), potentially limiting the broader applicability of these findings.
+Future research could address missing data through techniques like multiple imputation, and consider implementing a competing events analysis to provide a more comprehensive view of survival outcomes.
+Another limitation of this study is the slight deviations from the proportional hazards assumption in the Cox model for several covariates, including subtype, grade, radiation, chemotherapy, and surgery.
+Although the correlations between Schoenfeld residuals and ranked survival times were statistically significant, their small magnitudes suggest minimal practical impact.
+We proceeded with the Cox model, but these deviations should be interpreted with caution.
+Future research may consider using time-varying coefficients to address potential non-proportionality in hazard ratios.
+
+Furthermore, this study did not examine the combined effects of treatment modalities, such as how chemotherapy, radiation, and surgery work in synergy.
+Future research should focus on evaluating specific treatment combinations to optimize multimodal therapy for IBC patients.
+Lastly, while this study focused exclusively on IBC, comparing molecular subtypes in IBC patients to those in non-IBC patients could yield important insights into how subtype-specific outcomes differ between forms of breast cancer.
+Such comparisons could inform whether treatment strategies should differ by cancer type.
+
+# References

### 2024-12-05 code/analysis/table3.R

@@ -1,14 +1,22 @@
+###
 # Table 3: Survival for IBC by Clinicopathological Characteristics
+# MJ 2024-12-04
+# This script creates table 3 of the manuscript, presenting the median, 1-, 5-,
+# and 10- year survival for each covariate.
+###
 
+## Setup ####
 library(survival)
 library(gtsummary)
 library(dplyr)
 
-# load data
+# Load data
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-# 1, 5, and 10 year survival
+## Create table with 1, 5, and 10 year survival ###
+# Note that we create a filler column of 0-year survival to be replaced with
+# median survival later.
 t3 <- gtsummary::as_tibble(
   seer |>
   gtsummary::tbl_survfit(
@@ -32,9 +40,10 @@ t3 <- gtsummary::as_tibble(
   add_p()
 )
 
+# Assign cleaner column names
 colnames(t3) <- c("Characteristic","Median Survival (years)", "1 Year", "5 Year", "10 Year", "p-value")
 
-# create and assign more descriptive labels
+# Create and assign more descriptive labels
 t3$Characteristic = c(
   "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
   "Age", "20-49 years", "50-69 years", ">=70 years",
@@ -47,8 +56,7 @@ t3$Characteristic = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-# since we can't do both median survival and specific times,
-# create a seperate table with median survival
+## Create a separate table with median survival ####
 t3_med <- gtsummary::as_tibble(
   seer |>
     gtsummary::tbl_survfit(
@@ -72,17 +80,19 @@ t3_med <- gtsummary::as_tibble(
     )
 )
 
+# Assign cleaner column names
 colnames(t3_med) <- c("Characteristic","med_survival")
+# Ensure that median survival is numeric
 t3_med$med_survival <- as.numeric(t3_med$med_survival)
-
-# we want survival to be in years instead on months
+# Tranform the median survival from month to years
 t3_med$med_survival <- round(as.numeric(t3_med$med_survival / 12), 1)
 
-# overwrite the first stats column in table 2 with median survival information
+## Replace the filler column in the first table with median survival ####
 t3$`Median Survival (years)` <- t3_med$med_survival
 
[…]

### 2024-12-05 code/analysis/table4.R

@@ -1,10 +1,20 @@
+###
 # Table 4: Cox Proportional Hazards Model Regression Results
+# MJ 2024-12-04
+# This script creates table 4 of the manuscript, presenting the crude and 
+# adjusted cox proportional hazards models.
+###
 
-# load data
+## Setup ####
+library(survival)
+library(dplyr)
+
+# Load data
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-# first univariate table
+
+## Create the univariate coxph model and table ####
 t4_uv <- gtsummary::tbl_uvregression(
   data = seer,
   method = coxph,
@@ -26,9 +36,7 @@ t4_uv <- gtsummary::tbl_uvregression(
   hide_n = TRUE,
   )
 
-#colnames(t4_uv) <- c("Characteristic","HR", "95% CI", "p-value")
-
-# create and assign more descriptive labels
+# Create and assign more descriptive labels
 t4_uv$table_body$label[1:36] = c(
   "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
   "Age", "20-49 years", "50-69 years", ">=70 years",
@@ -41,7 +49,7 @@ t4_uv$table_body$label[1:36] = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-# next multivariable table
+## Create the multivariate coxph model and table ####
 t4_mv <- coxph(
     Surv(survival_months, survival_status) ~ 
       subtype + 
@@ -72,9 +80,7 @@ t4_mv <- coxph(
       )
     )
 
-#colnames(t4_mv) <- c("Characteristic","HR", "95% CI", "p-value")
-
-# create and assign more descriptive labels
+# Create and assign more descriptive labels
 t4_mv$table_body$label = c(
   "Subtype", "Luminal A", "Luminal B", "Triple Negative", "HER2 Positive",
   "Age", "20-49 years", "50-69 years", ">=70 years",
@@ -87,8 +93,8 @@ t4_mv$table_body$label = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-# Stick them together to make table 3
-t4 <- gtsummary::as_tibble(
+## Stick the univariate and multivariate tables together to make table 4 ####
+t4_merge <- gtsummary::as_tibble(
   gtsummary::tbl_merge(
     list(t4_uv, t4_mv),
     tab_spanner = c("Crude", "Adjusted")
[…]

### 2024-12-05 products/manuscript/manuscript.qmd
a few edits

@@
 Inflammatory breast cancer (IBC) is a rare and aggressive form of breast cancer characterized by rapid progression and poor prognosis.
-Molecular subtypes of breast cancer, such as Luminal A, Luminal B, HER2 Positive, and Triple Negative, have been associated with varying survival outcomes.
+The molecular subtypes of breast cancer, Luminal A, Luminal B, HER2 Positive, and Triple Negative, have been associated with varying survival outcomes.
 However, the impact of these subtypes specifically on IBC mortality remains underexplored.
@@
 
-Because IBC is rare, the biological characteristics of IBC are not commonly reported and the molecular alterations that result in poor prognosis are not well understood.
+Because IBC is rare, its biological characteristics are not commonly reported and the molecular alterations that result in poor prognosis are not well understood.
 However, hormone receptors (HR) and human epidermal growth factor receptor-2 (HER2), which define the molecular status of IBC through immunohistochemistry (IHC), are fundamental markers used to demonstrate molecular features, predict prognosis, and optimize therapeutic regimens [@li2017outcomes].
@@
 The proportional hazards assumption was assessed visually with log minus log plots and statistically by calculating the Pearson's product moment correlation coefficient of the Schoenfeld residuals and ranked survival times.
-After determining the assumption of proportional hazard was not violated, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
-We obtained hazard ratios (HR) and 95% confidence intervals (CI) for each covariate.
+After determining the assumption of proportional hazards was not violated, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
+We obtained hazard ratios (HR), 95% confidence intervals (CI), and p-values for each covariate.
 All statistical analyses were performed using version 3.7-0 of the survival package in R version 4.3.3.
@@
 Moreover, Luminal B and HER2 Positive subtypes appear to affect younger age groups more frequently than Luminal A, with 85.4% and 84.5% of cases, respectively, occurring in patients under 69 years, compared to 76.5% for Luminal A.
-A higher proportion of Black patients is observed in the Triple-Negative group compared to other subtypes.
+A higher proportion of Black patients is observed in the Triple Negative group compared to other subtypes.
 Comparing marital status reveals that the Triple Negative subtype affects the highest percentage of unmarried women.
@@
 
-\clearpage
-
 # Discussion
 
-In this large population-based cohort of adult women diagnosed with IBC, we found improved survival in Luminal B molecular subtype and poorer in Triple Negative.
+In this large population-based cohort of adult women diagnosed with IBC, we found better survival in Luminal B molecular subtype patients and poorer in Triple Negative.
 These findings are consistent with previous literature, which similarly identified Triple Negative IBC as a particularly aggressive form of breast cancer with poorer prognoses compared to hormone receptor-positive subtypes [@li2017outcomes, @pan2019nomogram, @menta2018inflammatory].
@@
 Furthermore, unmarried women experienced worse outcomes, consistent with studies suggesting that social support and marital status are important predictors of cancer survival [@osborne2005influence, @martinez2017prognostic].
+This could be related to the higher incidence of Triple Negative subtype observed in unmarried patients.
+Additionally, prior studies have found that in the U.S., Black women experience lower marriage rates compared to other racial and ethnic groups and face higher rates of marital instability, with divorce rates exceeding those of White and Hispanic women [@raley2015growing].
 

### 2024-12-05 code/analysis/flowchart.R

@@ -1,12 +1,17 @@
+###
 # Figure 1: Flowchart of IBC patients screening on SEER database
+# MJ 2024-12-05
+# This script created the Figure 1 flowchart describing our study population.
+###
 
+## Setup =======================================================================
 library(Gmisc, quietly = TRUE)
 library(glue)
 library(htmlTable)
 library(grid)
 library(magrittr)
 
-
+## Create the elgible, included, and excluded boxes of the flowchart ===========
 eligible <- boxGrob(glue("Adult female patients diagnosed with \ninflammatory breast carcinoma between 2010-2021",
                          "N = {pop}",
                          pop = txtInt(5969),
@@ -28,32 +33,34 @@ excluded <- boxGrob(glue("Excluded (n = {tot}):",
                          .sep = "\n"),
                     just = "left")
 
+## Set up the png file where the flowchart will be saved =======================
 png("results/figures/flowchart.png", width = 700, height = 500)
 
 grid.newpage()
-
+# Established the vertical spread of the figure
 vert <- spreadVertical(eligible = eligible,
                        included = included,
                        .from = .9,
                        .to = .1,
                        .type = "center")
-
+# Establishes the excluded box
 excluded <- moveBox(excluded,
                     x = .8,
                     y = coords(vert$included)$top + 
                       distance(vert$eligible, 
                                vert$included, half = TRUE, center = FALSE))
-
+# Draws the arrows on the flowchart
 for (i in 1:(length(vert) - 1)) {
   connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
     print
 }
-
 connectGrob(vert$eligible, excluded, type = "L")
-grid.rect(gp = gpar(col = "black", fill = NA, lwd = 2))  # Adjust color and line width as needed
+grid.rect(gp = gpar(col = "black", fill = NA, lwd = 2))
 
-# Print boxes
+# Prints boxes
 vert
 excluded
 
 dev.off()
+
+## End of script ===============================================================

### 2024-12-05 code/analysis/loglog-plots.R

@@ -1,5 +1,10 @@
-# Figure 2: Log(-log) plots to check Proportional Hazards assumption
+###
+# Figure 2: Log minus plots to check Proportional Hazards assumption
+# MJ 2024-12-05
+# This script creates Figure 2 log minus log plots
+###
 
+## Setup =======================================================================
 library(gridExtra)
 library(survminer)
 library(survival)
@@ -8,7 +13,7 @@ library(survival)
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-# create each plot
+## Create each plot ============================================================
 plot_subtype <- ggsurvplot(
   survfit(Surv(time = survival_months, event = survival_status) ~ subtype), 
   data = seer, 
@@ -19,7 +24,7 @@ plot_subtype <- ggsurvplot(
   legend.title = "Subtype:")
 
 plot_agecat <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ agecat), 
+  survfit(Surv(time = survival_months, event = survival_status) ~ age), 
   data = seer, 
   fun="cloglog",
   ylab = "log(-log(S(t))",
@@ -28,7 +33,7 @@ plot_agecat <- ggsurvplot(
   legend.title = "Age:")
 
 plot_race <- ggsurvplot(
-  survfit(Surv(time = survival_months, event = survival_status) ~ racecat), 
+  survfit(Surv(time = survival_months, event = survival_status) ~ race), 
   data = seer, 
   fun="cloglog",
   ylab = "log(-log(S(t))",
@@ -99,8 +104,7 @@ plot_surgery <- ggsurvplot(
   legend.labs = c("No/Unknown", "Yes"),
   legend.title = "Surgery:")
 
-# arrange on one grid and save as png
-
+## Arrange on one grid and save as png =========================================
 png("results/figures/loglog_plot.png", width = 900, height = 1100)
 
 grid.arrange(plot_subtype$plot, plot_agecat$plot, plot_race$plot, plot_marriage$plot, plot_n_stage$plot,
@@ -108,3 +112,5 @@ grid.arrange(plot_subtype$plot, plot_agecat$plot, plot_race$plot, plot_marriage$
              ncol = 2, nrow = 5)
 
 dev.off()
+
+## End of script ===============================================================

### 2024-12-05 code/analysis/table1.R

@@ -1,7 +1,11 @@
+###
 # Table 1: Clinicopathological Characteristics by Breast Cancer Subtype
-# (unadjusted associations between covariates and exposure)
+# MJ 2024-12-05
+# This script creates Table 1 of the manuscript. That is, the unadjusted 
+# associations between covariates and exposure (subtype)
+###
 
-# Declare dependencies
+## Setup =======================================================================
 library(here)
 library(readr)
 library(gt)
@@ -14,7 +18,7 @@ library(stringr)
 seer <- readRDS("data/processed/seer_tab.rds")
 attach(seer)
 
-# create table 1
+## Create table 1 ==============================================================
 t1 <- gtsummary::as_tibble(
   seer |>
   gtsummary::tbl_summary(
@@ -51,8 +55,11 @@ t1 <- gtsummary::as_tibble(
 # Remove Markdown bold (`**`) from the tibble text
 colnames(t1) <- stringr::str_remove_all(colnames(t1), "\\*\\*")
 
-# Convert the data frame to a flextable
+# Convert the tibble to a flextable
 t1_flex <- flextable::flextable(t1)
-t1_flex <- flextable::fontsize(t1_flex, size = 7.5, part = "all")
-# Save the flextable object as RDS
+t1_flex <- flextable::fontsize(t1_flex, size = 8.5, part = "all")
+
+## Save as RDS ================================================================
 saveRDS(t1_flex, file = here::here("results/tables", "t1.rds"))
+
+# End of script ================================================================

### 2024-12-05 code/analysis/table2.R

@@ -1,5 +1,12 @@
-# Table 2: Schoenfeld residual and ranked survival time correlations 
+###
+# Table 2: Schoenfeld residual and ranked survival time correlations
+# MJ 2024-12-05
+# This script creates table 2 of the manuscript. That is, the correlations
+# between the schoenfeld residuals from the full coxph model and the ranked
+# survival times.
+###
 
+## Setup =======================================================================
 library(survival)
 library(dplyr)
 library(gtsummary)
@@ -9,7 +16,7 @@ library(gt)
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-# fit the full cox proportional hazards model
+## Fit the full cox proportional hazards model =================================
 model <- coxph(
   Surv(survival_months, survival_status) ~ 
     relevel(factor(subtype), ref = '0') +
@@ -25,10 +32,10 @@ model <- coxph(
   data = seer, 
   method = "breslow")
 
-# extract Schoenfeld residuals
+# Extract Schoenfeld residuals
 resid <- cox.zph(model, terms = FALSE)
 
-# clean up to present in table
+# Clean up the covariate names to present in table
 covariates <- c("Luminal B", "Triple Negative", "HER2 Positive", 
                 "50-69 years", ">=70 years",
                 "Black", "Other",
@@ -40,7 +47,7 @@ covariates <- c("Luminal B", "Triple Negative", "HER2 Positive",
                 "Chemotherapy: Yes",
                 "Surgery: Yes")
 
-# get the correlations between residuals and ranked survival times
+## Calculate the correlations between residuals and ranked survival times ======
 cor <- data.frame(
   Variable = covariates,
   Correlation = apply(resid$y, 2, \(x) round(cor.test(rank(resid$time), x)$estimate, 3)),
@@ -50,11 +57,14 @@ cor <- data.frame(
   })
 )
 
+# Turn into a tibble
 t2 <- gtsummary::as_tibble(cor)
 
-# Convert to a flextable
+# Convert to flextable
 t2_flex <- flextable::flextable(t2)
 t2_flex <- flextable::fontsize(t2_flex, size = 7.5, part = "all")
-# Save the flextable object as RDS
+
+## Save as RDS =================================================================
 saveRDS(t2_flex, file = here::here("results/tables", "t2.rds"))
 
+## End of script ===============================================================

### 2024-12-05 code/analysis/table3.R

@@ -5,7 +5,7 @@
 # and 10- year survival for each covariate.
 ###
 
-## Setup ####
+## Setup =======================================================================
 library(survival)
 library(gtsummary)
 library(dplyr)
@@ -14,7 +14,7 @@ library(dplyr)
 seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
-## Create table with 1, 5, and 10 year survival ###
+## Create table with 1, 5, and 10 year survival ================================
 # Note that we create a filler column of 0-year survival to be replaced with
 # median survival later.
 t3 <- gtsummary::as_tibble(
@@ -56,7 +56,7 @@ t3$Characteristic = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-## Create a separate table with median survival ####
+## Create a separate table with median survival ================================
 t3_med <- gtsummary::as_tibble(
   seer |>
     gtsummary::tbl_survfit(
@@ -87,12 +87,14 @@ t3_med$med_survival <- as.numeric(t3_med$med_survival)
 # Tranform the median survival from month to years
 t3_med$med_survival <- round(as.numeric(t3_med$med_survival / 12), 1)
 
-## Replace the filler column in the first table with median survival ####
+## Replace the filler column in the first table with median survival ===========
 t3$`Median Survival (years)` <- t3_med$med_survival
 
-## Convert the data frame to a flextable ####
+## Convert the data frame to a flextable =======================================
 t3_flex <- flextable::flextable(t3)
-t3_font <- flextable::fontsize(t3_flex, size = 7.5, part = "all")
+t3_font <- flextable::fontsize(t3_flex, size = 9, part = "all")
 
-## Save the flextable object as RDS ####
-saveRDS(t3_flex, file = here::here("results/tables", "t3.rds"))
+## Save as RDS =================================================================
+saveRDS(t3_font, file = here::here("results/tables", "t3.rds"))
+
+## End of script ===============================================================

### 2024-12-05 code/analysis/table4.R

@@ -5,7 +5,7 @@
 # adjusted cox proportional hazards models.
 ###
 
-## Setup ####
+## Setup =======================================================================
 library(survival)
 library(dplyr)
 
@@ -14,13 +14,14 @@ seer <- readRDS("data/processed/seer.rds")
 attach(seer)
 
 
-## Create the univariate coxph model and table ####
+## Create the univariate coxph model and table =================================
 t4_uv <- gtsummary::tbl_uvregression(
   data = seer,
   method = coxph,
   y = Surv(survival_months, survival_status),
   exponentiate = TRUE,
-  include = c(subtype, age, race, marriage_status, n_stage, m_stage, grade, radiation, chemotherapy, surgery),
+  include = c(subtype, age, race, marriage_status, n_stage, m_stage, grade, 
+              radiation, chemotherapy, surgery),
   label = list(
     subtype ~ "Subtype", 
     age ~ "Age",
@@ -49,7 +50,7 @@ t4_uv$table_body$label[1:36] = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-## Create the multivariate coxph model and table ####
+## Create the multivariate coxph model and table ===============================
 t4_mv <- coxph(
     Surv(survival_months, survival_status) ~ 
       subtype + 
@@ -93,7 +94,7 @@ t4_mv$table_body$label = c(
   "Chemotherapy", "No/Unknown", "Yes",
   "Surgery", "No/Unknown", "Yes")
 
-## Stick the univariate and multivariate tables together to make table 4 ####
+## Stick the univariate and multivariate tables together to make table 4 =======
 t4_merge <- gtsummary::as_tibble(
   gtsummary::tbl_merge(
     list(t4_uv, t4_mv),
@@ -120,7 +121,9 @@ og_names <- c("Characteristic", "HR", "95% CI", "p-value",
               "aHR", "95% CI", "p-value")
 t4_og <- flextable::set_header_labels(t4_flex, 
                                    values = setNames(og_names, temp_colnames))
-t4 <- flextable::fontsize(t4_og, size = 7.5, part = "all")
+t4 <- flextable::fontsize(t4_og, size = 9, part = "all")
 
-## Save as rds ####
+## Save as rds =================================================================
 saveRDS(t4, file = here::here("results/tables", "t4.rds"))
+
+## End of script ===============================================================

### 2024-12-05 code/processing/processing-script.R

@@ -1,4 +1,10 @@
+###
+# Processing script
+# MJ 2024-12-05
+# This script cleans and processes the raw data pulled from the SEER database.
+###
 
+## Setup ======================================================================
 library(dplyr)
 library(stringr)
 library(psych)
@@ -6,10 +12,10 @@ library(gmodels)
 library(survival)
 library(survminer)
 
-# load data
+# Load data
 seer1 <- read.csv("data/raw/caselist_2024.11.csv", header = TRUE)
 
-# column rename
+## Rename the columns for easier workflow ======================================
 seer_name <- seer1 %>%
   dplyr::rename(
     age = Age.recode.with..1.year.olds,
@@ -39,12 +45,15 @@ seer_name <- seer1 %>%
     idc_type = Histologic.Type.ICD.O.3
   )
 
+## Filter the data by exclusion criteria =======================================
+# Restrict IDC code to 8500 - 8549
 seer_idc <- seer_name %>%
   filter(
     idc_type %in% seq(8500, 8549) # 481 indiv exclude, 5986 remain
   )
 
-# filter by exclusion criteria
+# Restrict to adult females who survival longer than 1 month and whose cause of
+# death, if dead, is attributable to IBC 
 seer_filter <- seer_idc %>%
   filter(
     age != "00 years",
@@ -57,7 +66,8 @@ seer_filter <- seer_idc %>%
     survival_cause_other != "Dead (attributable to causes other than this cancer dx)", # 487 indiv exclude, 5317 remain
   )
 
-# clean up the variables
+## Clean up existing variable categories and merge variables that were measured
+# differently across years =====================================================
 seer_clean <- seer_filter %>%
   mutate(
     # assign race as White, Black, or Other
@@ -71,7 +81,6 @@ seer_clean <- seer_filter %>%
       marriage_status == "Married (including common law)" ~ "Yes",
       TRUE ~ "No"
     ),
-    
     # assign blank tnm to NA
     tnm2015 = case_when(
       tnm2015 == "Blank(s)" ~ NA_character_,
@@ -85,7 +94,6 @@ seer_clean <- seer_filter %>%
       tnm2018 == "Blank(s)" ~ NA_character_,
       TRUE ~ tnm2018
     ),
-    
     # merge tnm to ensure cross-validation
     tnm = coalesce(tnm2015, tnm2017, tnm2018),
     
[…]

### 2024-12-06 products/manuscript/manuscript.qmd
another round of proofreading

@@
 
-For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N (node) stage (N0, N1, N2, N3), M (metastasis) stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), and survival months.
+For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \>=70), race (white, black, other), marital status (unmarried, married), N (node) stage (N0, N1, N2, N3), M (metastasis) stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival status (Alive, Dead), and survival months.
 Breast cancer subtypes were classified as Luminal A (HR+/HER2-), Luminal B (HR+/HER+), HER2 Positive (HR-/HER2+), and Triple Negative.
@@
 Clinicopathological characteristics were compared between IBC subtype using Pearson's Chi-squared test.
-Median, 1-, 5-, and 10-year survival for IBC is estimated for each variable using the Kaplain-Meier method and compared using the log-rank test in the subgroups.
+Median, 1-, 5-, and 10-year survival for IBC is estimated for each variable using the Kaplain-Meier method and compared using the log-rank test within the subgroups.
 The proportional hazards assumption was assessed visually with log minus log plots and statistically by calculating the Pearson's product moment correlation coefficient of the Schoenfeld residuals and ranked survival times.
-After determining the assumption of proportional hazards was not violated, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
+After determining that the assumption of proportional hazards was not violated, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
 We obtained hazard ratios (HR), 95% confidence intervals (CI), and p-values for each covariate.
@@
 A total of 5969 adult female IBC cases were eligible from 2010-2021 in the SEER database for our study population.
-Of these, 1829 were excluded; 165 had a survival time less than 1 month, 487 had an unknown or competing cause of death, and 1177 had incomplete information.
+Of these, 1829 were excluded; 165 had a survival time of less than 1 month, 487 had an unknown or competing cause of death, and 1177 had incomplete information.
 The remaining 4140 were included in our analysis, as shown in @fig-flowchart.
@@
 
-The proportional hazards assumption was assessed for each covariate visually using log minus log plots, as shown in @fig-loglog.
+The proportional hazards assumption was assessed for each covariate visually using log minus log plots, shown in @fig-loglog.
 
@@
 
-Upon inspection, slight deviations from parallelism were observed for the covariates subtype, grade, radiation, chemotherapy, and surgery.
+Upon inspection, slight deviations from parallelism were observed for subtype, grade, radiation, chemotherapy, and surgery.
 Therefore, Schoenfeld residuals were extracted from the full Cox proportional hazards model.
@@
 Survival differences are also pronounced by distant metastasis, with M1 stage associated with worse outcomes.
-Additionally, higher cancer grades correlate with poorer survival.
+Higher cancer grades also correlate with poorer survival.
 In terms of treatment, patients who receive radiation, chemotherapy, or surgery exhibit better survival outcomes across all measurements.
@@
 
-We conducted univariate and multivariate analyses based on the data in @tbl-table3. The results of the Cox proportional hazards regression models are presented in @tbl-table4.
+We conducted univariate and multivariate analyses based on the data in @tbl-table3.
+The results of the Cox proportional hazards regression models are presented in @tbl-table4.
+Patients who have an end survival status of "Alive" are considered censored.
+
+## Univariate Analysis
+
+In the univariate analysis, subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were all significantly associated with IBC-specific survival.
+These variables were therefore included in the multivariate analysis.
+
+## Multivariate Analysis
+
+The multivariate analysis revealed that subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were independent predictors of IBC-specific survival.
+The Luminal B subtype was associated with better survival (aHR=0.72, 95%CI: 0.56-0.74, P\<0.001), whereas the Triple Negative subtype showed the poorest survival outcomes (aHR=2.51, 95%CI: 1.79-2.21, P\<0.001).
+In terms of treatment, our results found that the existing therapies are effective in improving survival.
+That is, radiation (aHR=0.76, 95%CI: 0.69-0.85, P\<0.001), chemotherapy (aHR=0.50, 95%CI: 0.44-0.57, P\<0.001), and surgery (aHR=0.59, 95%CI: 0.52-0.66, P\<0.001).
 
@@
 ```
-
-## Univariate Analysis
-
-In the univariate analysis, subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were all significantly associated with IBC-specific survival.
-These variables were therefore included in the multivariate analysis.
-
-## Multivariate Analysis
-
-The multivariate analysis revealed that subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were independent predictors of IBC-specific survival.
-The Luminal B subtype was associated with better survival (aHR=0.72, 95%CI: 0.56-0.74, P\<0.001), whereas the Triple Negative subtype showed the poorest survival outcomes (aHR=2.51, 95%CI: 1.79-2.21, P\<0.001).
-In terms of treatment, our results found that the existing therapies are effective in improving survival.
-That is, radiation (aHR=0.76, 95%CI: 0.69-0.85, P\<0.001), chemotherapy (aHR=0.50, 95%CI: 0.44-0.57, P\<0.001), and surgery (aHR=0.59, 95%CI: 0.52-0.66, P\<0.001).
 
@@
 
+# Conclusions
+
+This study identified molecular subtype as a crucial predictor of cancer-specific survival in women with inflammatory breast cancer, with the Triple Negative subtype associated with the poorest outcomes and Luminal B showing the best prognosis.
+Additionally, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were all found to be independent prognostic factors for survival.
+These findings emphasize the need for personalized therapeutic strategies, particularly for patients with Triple Negative IBC, and highlight the importance of addressing the racial and social inequities in IBC survival outcomes.
+
 # References

### 2025-03-20 products/manuscript/manuscript.qmd
fix: manuscript typo

@@
 The proportional hazards assumption was assessed visually with log minus log plots and statistically by calculating the Pearson's product moment correlation coefficient of the Schoenfeld residuals and ranked survival times.
-After determining that the assumption of proportional hazards was not violated, univariate and multivariate Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
+After determining that the assumption of proportional hazards was not violated, univariable and multivariable Cox proportional hazards regressions were used to model the relationship between potential covariates and IBC-specific survival.
 We obtained hazard ratios (HR), 95% confidence intervals (CI), and p-values for each covariate.
@@
 
-We conducted univariate and multivariate analyses based on the data in @tbl-table3.
+We conducted univariable and multivariable analyses based on the data in @tbl-table3.
 The results of the Cox proportional hazards regression models are presented in @tbl-table4.
@@
 
-## Univariate Analysis
+## Univariable Analysis
 
-In the univariate analysis, subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were all significantly associated with IBC-specific survival.
-These variables were therefore included in the multivariate analysis.
+In the univariable analysis, subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were all significantly associated with IBC-specific survival.
+These variables were therefore included in the multivariable analysis.
 
-## Multivariate Analysis
+## Multivariable Analysis
 
-The multivariate analysis revealed that subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were independent predictors of IBC-specific survival.
+The multivariable analysis revealed that subtype, age, race, marital status, N stage, M stage, grade, radiation, chemotherapy, and surgery were independent predictors of IBC-specific survival.
 The Luminal B subtype was associated with better survival (aHR=0.72, 95%CI: 0.56-0.74, P\<0.001), whereas the Triple Negative subtype showed the poorest survival outcomes (aHR=2.51, 95%CI: 1.79-2.21, P\<0.001).
@@
 
-This study has several strengths, including the use of a large, diverse patient population and robust multivariate analyses that accounted for key clinicopathological factors.
+This study has several strengths, including the use of a large, diverse patient population and robust multivariable analyses that accounted for key clinicopathological factors.
 However, several limitations should be acknowledged.
